### 2021-09-15 instrument.tex
Update

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Helvetica}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Light Exposure Behavior Assessment (LEBA): Long and Short Form},
  pdfauthor={Mushfiqul Anwar Siraji, Rafael Robert Lazar, Juliëtte van Duijnhoven, Luc Schlangen, Shamsul Haque, Vineetha Kalavally, Céline Vetter, Gena Glickman, Karin Smolders, \& Manuel Spitschan},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{LEBA}
\usepackage{csquotes}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{\emph{Light Exposure Behavior Assessment (LEBA)}: Long and Short Form}
\author{Mushfiqul Anwar Siraji\textsuperscript{}, Rafael Robert Lazar\textsuperscript{}, Juliëtte van Duijnhoven\textsuperscript{}, Luc Schlangen\textsuperscript{}, Shamsul Haque\textsuperscript{}, Vineetha Kalavally\textsuperscript{}, Céline Vetter\textsuperscript{}, Gena Glickman\textsuperscript{}, Karin Smolders\textsuperscript{}, \& Manuel Spitschan\textsuperscript{}}
\date{}

\affiliation{\phantom{0}}

\begin{document}
\maketitle

\textbf{Disclaimer}: This is a non-public version of LEBA and still a work in progress.

LEBA captures light exposure-related behaviours on a 5 point Likert type scale ranging from 1 to 5 (Never =1; Rarely= 2; Sometimes =3; Often = 4; Always =5). The score of each factor is calculated by the summation of scores of items belonging to the corresponding factor.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{1.00}}@{}}
\toprule
Long Form (23 Items) \\
\midrule
\endhead
\textbf{F1: Blue Filter} \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses indoors during the day. \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses outdoors during the day. \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses within 1 hour before attempting to fall asleep. \\
\textbf{F2: Natural Light} \\
I spend 30 minutes or less per day (in total) outside. \\
I spend between 1 and 3 hours per day (in total) outside. \\
I spend between 30 minutes and 1 hour per day (in total) outside. \\
I spend more than 3 hours per day (in total) outside. \\
I spend as much time outside as possible. \\
I go for a walk or exercise outside within 2 hours after waking up. \\
\textbf{F3: Smart Device} \\
I use my mobile phone within 1 hour before attempting to fall asleep. \\
I look at my mobile phone screen immediately after waking up. \\
I check my phone when I wake up at night. \\
I look at my smartwatch within 1 hour before attempting to fall asleep. \\
I look at my smartwatch when I wake up at night. \\
\textbf{F4: Sleep Environment} \\
I dim my mobile phone screen within 1 hour before attempting to fall asleep. \\
I use a blue-filter app on my computer screen within 1 hour before attempting to fall asleep. \\
I use as little light as possible when I get up during the night. \\
I dim my computer screen within 1 hour before attempting to fall asleep. \\
\textbf{F5: Electric Light} \\
I use tunable lights to create a healthy light environment. \\
I use LEDs to create a healthy light environment. \\
I use a desk lamp when I do focused work. \\
I use an alarm with a dawn simulation light. \\
I turn on the lights immediately after waking up. \\
\bottomrule
\end{longtable}

\newpage

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{1.00}}@{}}
\toprule
Short Form (17 Items) \\
\midrule
\endhead
\textbf{F1: Blue Filter} \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses indoors during the day. \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses outdoors during the day. \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses within 1 hour before attempting to fall asleep. \\
\textbf{F2: Natural Light} \\
I spend 30 minutes or less per day (in total) outside. \\
I spend between 1 and 3 hours per day (in total) outside. \\
I spend more than 3 hours per day (in total) outside. \\
I spend as much time outside as possible. \\
I go for a walk or exercise outside within 2 hours after waking up. \\
\textbf{F3: Smart Device} \\
I use my mobile phone within 1 hour before attempting to fall asleep. \\
I look at my mobile phone screen immediately after waking up. \\
I check my phone when I wake up at night. \\
\textbf{F4: Sleep Environment} \\
I dim my mobile phone screen within 1 hour before attempting to fall asleep. \\
I use a blue-filter app on my computer screen within 1 hour before attempting to fall asleep. \\
I dim my computer screen within 1 hour before attempting to fall asleep. \\
\textbf{F5: Electric Light} \\
I use tunable lights to create a healthy light environment. \\
I use LEDs to create a healthy light environment. \\
I use an alarm with a dawn simulation light. \\
\bottomrule
\end{longtable}

\end{document}