### 2021-10-22 Appendix A.Rmd
Factor name finalised

---
title: "Appendix A"
author: "Team LEBA"
date: "10/22/2021"
output: pdf_document
---

```{r setup, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, dpi=600)
set.seed(123)
par(family = "Helvetica")
```