### 2021-10-22 instrument.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
\textbf{Disclaimer}: This is a non-public version of LEBA (dated \today)
and still a work in progress. Please do not distribute!

LEBA captures light exposure-related behaviours on a 5 point Likert type
scale ranging from 1 to 5 (Never/Does not apply/I don't know = 1; Rarely
= 2; Sometimes = 3; Often = 4; Always = 5). The score of each factor is
calculated by the summation of scores of items belonging to the
corresponding factor.

\newpage

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{1.00}}@{}}
\toprule
Long Form (23 Items) \\
\midrule
\endhead
\textbf{F1: Wearing blue light filters} \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses indoors
during the day. \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses outdoors
during the day. \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses within 1
hour before attempting to fall asleep. \\
\textbf{F2: Spending time outdoors} \\
I spend 30 minutes or less per day (in total) outside. \\
I spend between 1 and 3 hours per day (in total) outside. \\
I spend between 30 minutes and 1 hour per day (in total) outside. \\
I spend more than 3 hours per day (in total) outside. \\
I spend as much time outside as possible. \\
I go for a walk or exercise outside within 2 hours after waking up. \\
\textbf{F3: Using phone and smartwatch in bed} \\
I use my mobile phone within 1 hour before attempting to fall asleep. \\
I look at my mobile phone screen immediately after waking up. \\
I check my phone when I wake up at night. \\
I look at my smartwatch within 1 hour before attempting to fall
asleep. \\
I look at my smartwatch when I wake up at night. \\
\textbf{F4: Using light before bedtime} \\
I dim my mobile phone screen within 1 hour before attempting to fall
asleep. \\
I use a blue-filter app on my computer screen within 1 hour before
attempting to fall asleep. \\
I use as little light as possible when I get up during the night. \\
I dim my computer screen within 1 hour before attempting to fall
asleep. \\
\textbf{F5: Using light in the morning and during daytime} \\
I use tunable lights to create a healthy light environment. \\
I use LEDs to create a healthy light environment. \\
I use a desk lamp when I do focused work. \\
I use an alarm with a dawn simulation light. \\
I turn on the lights immediately after waking up. \\
\bottomrule
\end{longtable}

\newpage

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{1.00}}@{}}
\toprule
Short Form (17 Items) \\
\midrule
\endhead
\textbf{F1: Wearing blue light filters} \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses indoors
during the day. \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses outdoors
during the day. \\
I wear blue-filtering, orange-tinted, and/or red-tinted glasses within 1
hour before attempting to fall asleep. \\
\textbf{F2: Spending time outdoors} \\
I spend 30 minutes or less per day (in total) outside. \\
I spend between 1 and 3 hours per day (in total) outside. \\
I spend more than 3 hours per day (in total) outside. \\
I spend as much time outside as possible. \\
I go for a walk or exercise outside within 2 hours after waking up. \\
\textbf{F3: Using phone and smartwatch in bed} \\
I use my mobile phone within 1 hour before attempting to fall asleep. \\
I look at my mobile phone screen immediately after waking up. \\
I check my phone when I wake up at night. \\
\textbf{F4: Using light before bedtime} \\
I dim my mobile phone screen within 1 hour before attempting to fall
asleep. \\
I use a blue-filter app on my computer screen within 1 hour before
attempting to fall asleep. \\
I dim my computer screen within 1 hour before attempting to fall
asleep. \\
\textbf{F5: Using light in the morning and during daytime} \\
I use tunable lights to create a healthy light environment. \\
I use LEDs to create a healthy light environment. \\
I use an alarm with a dawn simulation light. \\
\bottomrule
\end{longtable}
\end{appendix}