### 2021-10-29 instrument.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
\textbf{Disclaimer}: This is a non-public version of LEBA (dated \today)
and still a work in progress. Please do not distribute!

LEBA captures light exposure-related behaviours on a 5 point Likert type
scale ranging from 1 to 5 (Never/Does not apply/I don't know = 1; Rarely
= 2; Sometimes = 3; Often = 4; Always = 5). The score of each factor is
calculated by the summation of scores of items belonging to the
corresponding factor. The following instruction is given before
displaying the items: ``Please indicate how often you performed the
following behaviours in the past 4 weeks.''

\newpage

\blandscape

\hypertarget{leba-long-form-23-items}{%
\section{LEBA Long Form (23 Items)}\label{leba-long-form-23-items}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.05}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.27}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.34}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.08}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.11}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.07}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.08}}@{}}
\toprule
& Items & Never/Does not apply/I don't know & Rarely & Sometimes & Often
& Always \\
\midrule
\endhead
1 & I wear blue-filtering, orange-tinted, and/or red-tinted glasses
indoors during the day. & & & & & \\
2 & I wear blue-filtering, orange-tinted, and/or red-tinted glasses
outdoors during the day. & & & & & \\
3 & I wear blue-filtering, orange-tinted, and/or red-tinted glasses
within 1 hour before attempting to fall asleep. & & & & & \\
4 & I spend 30 minutes or less per day (in total) outside. & & & & & \\
5 & I spend between 1 and 3 hours per day (in total) outside. & & & &
& \\
6 & I spend between 30 minutes and 1 hour per day (in total) outside. &
& & & & \\
7 & I spend more than 3 hours per day (in total) outside. & & & & & \\
8 & I spend as much time outside as possible. & & & & & \\
9 & I go for a walk or exercise outside within 2 hours after waking up.
& & & & & \\
10 & I use my mobile phone within 1 hour before attempting to fall
asleep. & & & & & \\
11 & I look at my mobile phone screen immediately after waking up. & & &
& & \\
12 & I check my phone when I wake up at night. & & & & & \\
13 & I look at my smartwatch within 1 hour before attempting to fall
asleep. & & & & & \\
14 & I look at my smartwatch when I wake up at night. & & & & & \\
15 & I dim my mobile phone screen within 1 hour before attempting to
fall asleep. & & & & & \\
16 & I use a blue-filter app on my computer screen within 1 hour before
attempting to fall asleep. & & & & & \\
17 & I use as little light as possible when I get up during the night. &
& & & & \\
18 & I dim my computer screen within 1 hour before attempting to fall
asleep. & & & & & \\
19 & I use tunable lights to create a healthy light environment. & & & &
& \\
20 & I use LEDs to create a healthy light environment. & & & & & \\
21 & I use a desk lamp when I do focused work. & & & & & \\
22 & I use an alarm with a dawn simulation light. & & & & & \\
23 & I turn on the lights immediately after waking up. & & & & & \\
\bottomrule
\end{longtable}

\elandscape

\newpage

\blandscape

\hypertarget{latent-structure-reliability-and-structural-validity}{%
\subsection{Latent Structure, Reliability and Structural
Validity}\label{latent-structure-reliability-and-structural-validity}}

The long form of LEBA consists 23 items with five factors.

\begin{longtable}[]{@{}lll@{}}
\toprule
Factor names & Items & Reliability Coefficients \\
\midrule
\endhead
\textbf{F1: Wearing blue light filters} & 1-3 & .93 \\
\textbf{F2: Spending time outdoors} & 4-9 (Item 4 is reversed) & .80 \\
\textbf{F3: Using phone and smartwatch in bed} & 10-14 & .61 \\
\textbf{F4: Using light before bedtime} & 15-18 & .72 \\
\textbf{F5: Using light in the morning and during daytime} & 19-23 &
.45 \\
\bottomrule
\end{longtable}

LEBA -long form showed satisfactory structural validity (CFI =.97; TLI =
.96; RMSEA = .05{[}.04-.06, 90\% CI{]}; SRMR = .09).

\hypertarget{how-to-cite}{%
\subsection{How to cite:}\label{how-to-cite}}

\elandscape

\newpage

\blandscape

\hypertarget{leba-short-form-17-items}{%
\section{LEBA Short Form (17 Items)}\label{leba-short-form-17-items}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.05}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.27}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.34}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.08}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.11}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.07}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.08}}@{}}
\toprule
& Short Form (17 Items) & Never/Does not apply/I don't know & Rarely &
Sometimes & Often & Always \\
\midrule
\endhead
01 & I wear blue-filtering, orange-tinted, and/or red-tinted glasses
indoors during the day. & & & & & \\
02 & I wear blue-filtering, orange-tinted, and/or red-tinted glasses
outdoors during the day. & & & & & \\
03 & I wear blue-filtering, orange-tinted, and/or red-tinted glasses
within 1 hour before attempting to fall asleep. & & & & & \\
04 & I spend 30 minutes or less per day (in total) outside. & & & & & \\
05 & I spend between 1 and 3 hours per day (in total) outside. & & & &
& \\
06 & I spend more than 3 hours per day (in total) outside. & & & & & \\
07 & I spend as much time outside as possible. & & & & & \\
08 & I go for a walk or exercise outside within 2 hours after waking up.
& & & & & \\
09 & I use my mobile phone within 1 hour before attempting to fall
asleep. & & & & & \\
10 & I look at my mobile phone screen immediately after waking up. & & &
& & \\
11 & I check my phone when I wake up at night. & & & & & \\
12 & I dim my mobile phone screen within 1 hour before attempting to
fall asleep. & & & & & \\
13 & I use a blue-filter app on my computer screen within 1 hour before
attempting to fall asleep. & & & & & \\
14 & I dim my computer screen within 1 hour before attempting to fall
asleep. & & & & & \\
15 & I use tunable lights to create a healthy light environment. & & & &
& \\
16 & I use LEDs to create a healthy light environment. & & & & & \\
17 & I use an alarm with a dawn simulation light. & & & & & \\
\bottomrule
\end{longtable}

\elandscape

\newpage

\blandscape

\hypertarget{latent-structure-reliability-and-structural-validity-1}{%
\subsection{Latent Structure, Reliability and Structural
Validity}\label{latent-structure-reliability-and-structural-validity-1}}

The short form of LEBA consists 23 items with five factors.

\begin{longtable}[]{@{}ll@{}}
\toprule
Factor names & Items \\
\midrule
\endhead
\textbf{F1: Wearing blue light filters} & 1-3 \\
\textbf{F2: Spending time outdoors} & 4-8 (Item 4 is reversed) \\
\textbf{F3: Using phone and smartwatch in bed} & 9-11 \\
\textbf{F4: Using light before bedtime} & 12-14 \\
\textbf{F5: Using light in the morning and during daytime} & 15-17 \\
\bottomrule
\end{longtable}

\hypertarget{how-to-cite-1}{%
\subsection{How to cite:}\label{how-to-cite-1}}

\elandscape
\end{appendix}