### 2021-11-07 Supplymentary.Rmd
* supplementary analysis on EFA (Minimum residulas and six-factor solution) is added
Some minor formating here and there.

---
title: "Supplimentary Analysis"
output: pdf_document
---

```{r options, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, dpi=600)
set.seed(123)
par(family = "Helvetica")
```

```{r MinRes, include=F}
library(psych)
library(tabledown)
library(kableExtra)#without this pdf rendering will not happen

data <- readRDS("leba_2021-09-08.rds")

# Separating EFA and CFA samples with descriptive column
descriptives.data <- data

## Merge "0"s and "1"s into "1"s select,subset assign)
descriptives.data[ , 9:56 ][ descriptives.data[ , 9:56 ] == 0 ] <- 1
#EFA.descriptives <- descriptives.data[1:428,]
EFA.descriptives <- subset(descriptives.data, IncludedInEFA == "TRUE")
#CFA.descriptives <- descriptives.data[429:690, ]
CFA.descriptives<- subset(descriptives.data, IncludedInEFA == "FALSE")

#Separating the EFA and CFA(only items)
sem.data <- data[, 9:56] #EFA & CFA data
sem.data[ sem.data == 0] <- 1 #Merged "0"s and "1"s into "1"s

## renaming the column-header
prefix <- "item"
sufix <- c(1:48)
colnam <- paste(prefix,sufix, sep = "" )
colnames(sem.data) <- colnam

### EFA data
EFA.data <- sem.data[1:428, ]
#Polychoric Correlation matrix

correlations <- psych::polychoric(EFA.data, correct = 0)

#Supplementary EFA with Minres Extraction

fa.5F.min.1 <- fa(r=correlations$rho, nfactors = 5, fm= "minres",rotate ="varimax",
                residuals = TRUE, SMC = TRUE, n.obs =428)
EE <- print(fa.5F.min.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.5F.min.1 <- dplyr::select(EFA.data ,
                                    -c( item20, item19, item5, item31, item44, item24, item43, item39, item22, item2, item18, item48, item42, item29, item6, item15, item23, item28, item14))

correlations.red.5F.min.1 <- polychoric(reduced.model.5F.min.1,correct = 0)

fa.5F.min.2 <- fa(r=correlations.red.5F.min.1$rho, nfactors = 5, fm= "minres",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)

FF <- print(fa.5F.min.2, cut = .3, digits = 3, sort = TRUE)

reduced.model.5F.min.2 <- dplyr::select(EFA.data, -c( item20, item19, item5, item31, item44, item24, item43, item39, item22, item2, item18, item48, item42, item29, item6, item15, item23, item28, item14, item34, item47, item21, item13))

correlations.red.5F.min.2 <- polychoric(reduced.model.5F.min.2, correct = 0)

fa.5F.min.3 <- fa(r=correlations.red.5F.min.2$rho, nfactors = 5, fm= "minres",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)

GG <- print(fa.5F.min.3, cut = .3, digits = 3, sort = TRUE)

minresTab <- tabledown::fac.tab(fa.5F.min.3, cut =.3,complexity = F)
```

```{r MinResTab}
papaja::apa_table(minresTab, align = "c", caption = "Factor loadings and communality of the retained items(Minmum Residual)", note = "Only loading higher than .30 is reported", font_size = "small")

```

```{r sixFac, include=F}
# EFA with 6 factor (rejected)

#correlations <- polychoric(EFA.factor, correct = 0)

fa.6F.1 <- fa(r=correlations$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)

II <- print(fa.6F.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.6F.1 <- dplyr::select(EFA.data ,
                                    -c(item34, item44, item48, item43, item39, item22, item2, item20, item21, item24,item30, item14, item28, item6, item15, item23))

correlations.red.6F.1 <- polychoric(reduced.model.6F.1,correct = 0)

fa.6F.2 <- fa(r=correlations.red.6F.1$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)

JJ <- print(fa.6F.2, cut = .3, digits = 3, sort = TRUE)

reduced.model.6F.2 <- dplyr::select(EFA.data, -c(item34, item44, item48, item43, item39, item22, item2, item20, item21, item24,item30, item14, item28, item6, item15, item23, item18, item42, item47, item41))

correlations.red.6F.2 <- polychoric(reduced.model.6F.2,correct = 0)

fa.6F.3 <- fa(r=correlations.red.6F.2$rho, nfactors = 6, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428, max.iter = 500)

KK <- print(fa.6F.3, cut = .3, digits = 3, sort = TRUE)

# Residuals
residual.6F =residuals(fa.6F.3 , diag = FALSE, na.rm =T)
#Count of numbers of residuals>.05
BigRRR= sum(residual.6F>abs(.05), na.rm =T)
print(BigRRR)

#Total number of off diagonal elements in the data matrix
totRRR = length(reduced.model.6F.2)*(length(reduced.model.6F.2)-1)/2

# Proportion of off-diagonal elements >.10
sumRRR <- sum(BigRRR/totRRR*100)

hist(residual.6F, main = NULL, xlab = "Six factor Solution: Residuals",ylim=c(0, 400))
abline(h =65, lwd = 2, lty = 2)
abline(v =.05,lwd = 2, lty = 2)
sixFac <- tabledown::fac.tab(fa.6F.3, cut =.3,complexity = F)
```

```{r sixFacTab}
papaja::apa_table(sixFac, align = "c", caption = "Factor loadings and communality of the retained items(six factor)", note = "Only loading higher than .30 is reported", font_size = "small")
```