### 2022-01-02 appendix-1.Rmd
Duplicate chunk error fixed

```{r factormp, warning=F}
map <- readr::read_csv("Table_raw/map_stat")
map <- map[,-1]
papaja::apa_table(map,align = "c", caption = "MAP Statistics",  font_size = "small")
```

```{r F1IF, echo=FALSE, fig.align='center', fig.cap='Item information curve of LEBA F1', out.height='160%', out.width='100%'}
knitr::include_graphics('Manuscript.figures/F1testinfo.png')
```

```{r F2IF, echo=FALSE, fig.align='center', fig.cap='Item information curve of LEBA F1', out.height='160%', out.width='100%'}
knitr::include_graphics('Manuscript.figures/F2testinfo.png')
```

```{r F3IF, echo=FALSE, fig.align='center', fig.cap='Item information curve of LEBA F1', out.height='160%', out.width='100%'}
knitr::include_graphics('Manuscript.figures/F3testinfo.png')
```

```{r F4IF, echo=FALSE, fig.align='center', fig.cap='Item information curve of LEBA F1', out.height='160%', out.width='100%'}
knitr::include_graphics('Manuscript.figures/F4testinfo.png')
```

```{r F5IF, echo=FALSE, fig.align='center', fig.cap='Item information curve of LEBA F1', out.height='160%', out.width='100%'}
knitr::include_graphics('Manuscript.figures/F5testinfo.png')
```