### 2022-01-10 s1.Rmd

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : 
  papaja::apa6_pdf:
    keep_tex: true
    latex_engine: xelatex

appendix:
  - "appendix-1.Rmd"
mainfont: Arial
---
```{r library, include=FALSE}

library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot) #devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
library(semTools)
library(MOTE)
library(car)#reversecoding 
library(psych)
library(dlookr)
library(plyr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(DiagrammeR) #devtools::install_github('rich-iannone/DiagrammeR')
library(DiagrammeRsvg) #For DiagrammeR
library(rsvg) #For DiagrammeR
library(ggcorrplot)
library(semTable)
library(magick)
library(mirt)
library(gtsummary)
library(gt)
library(gtExtras)
library(likert)
library(VIM) #Missing data
library(kutils)
library(simsem)
library(tabledown)
library(readxl)
library(koRpus)# stable release
library(koRpus.lang.en)
r_refs("references.bib")
```

```{r setup, include = FALSE}
library("papaja")
r_refs("references.bib")
```

(ref:foo) Factor Identification Methods (A) Parallel analysis drew 500 sets of random data set (n=428 each) and adjusted the mean eigenvalues for sample bias. In x-axis we plotted the number of possible factors and in y-axis we plotted the mean adjusted eigenvalues. The First inflection point of the plot indicated the optimal number of factors were six. (B) In the scree plot possible number of factors are plotted against obtained eigenvalues. Scree plot suggested six factors. (C) Hull method we plotted the number of possible factors in the x-axis and the CAF index(CAF: Common Part Accounted for-an indicator of model fit) in the y-axis. Hull method indicated 5 factors were required to balance the model fit and number of parameters.

```{r foo, fig.cap='(ref:foo)',  warning=FALSE, results= 'asis',fig.align='center'}

knitr::include_graphics('Manuscript.figures/factors.png', dpi =800)

```

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup