### 2022-07-07 manuscript.Rmd
Adjusting phrasing, grammar, and spelling  in the analysis part of the manuscript.

---
title             : "_Light Exposure Behaviour Assessment (LEBA)_: Development of a novel instrument to capture light exposure-related behaviours"
shorttitle        : "LEBA"

author: 
  - name          : "Mushfiqul Anwar Siraji"
    affiliation   : "1, *"
    email         : "mushfiqul.anwarsiraji@monash.edu"
    role:         
      - Formal Analysis
      - Visualization
      - Writing – original draft
      - Writing – review & editing

  - name          : "Rafael Robert Lazar"
    affiliation   : "2, 3, *"
    role:
      - Data curation
      - Investigation
      - Project administration
      - Visualization
      - Writing – original draft
      - Writing – review & editing

  - name          : "Juliëtte	van Duijnhoven"
    affiliation   : "4, 5"
    email         : "j.v.duijnhoven1@tue.nl"
    role:         
      - Conceptualization
      - Methodology
      - Investigation
      - Writing – review & editing

  - name          : "Luc Schlangen"
    affiliation   : "5, 6"
    email         : "l.j.m.schlangen@tue.nl"
    role:         
      - Conceptualization
      - Methodology
      - Investigation
      - Writing – review & editing

  - name          : "Shamsul Haque"
    affiliation   : "1"
    email         : "shamsul@monash.edu"
    role:         
      - Conceptualization
      - Supervision
      - Writing – review & editing

  - name          : "Vineetha Kalavally"
    affiliation   : "7"
    email         : "vineetha@monash.edu"
    role:         
      - Supervision
      - Writing – review & editing

  - name          : "Céline Vetter"
    affiliation   : "8, 9"
    email         : "celine.vetter@colorado.edu"
    role:         
      - Conceptualization
      - Writing – review & editing

  - name          : "Gena Glickman"
    affiliation   : "10"
    email         : "gena.glickman@usuhs.edu"
    role:         
      - Conceptualization
      - Methodology
      - Writing – review & editing

  - name          : "Karin Smolders"
    affiliation   : "5,6"
    email         : "k.c.h.j.smolders@tue.nl"
    role:         
      - Conceptualization
      - Methodology
      - Writing – review & editing

  - name          : "Manuel Spitschan"
    corresponding : yes
    affiliation   : "11, 12, 13"
    email         : "manuel.spitschan@tum.de "
    role:         
      - Conceptualization
      - Data curation
      - Investigation
      - Project administration
      - Visualization
      - Methodology
      - Writing – original draft
      - Writing – review & editing

affiliation:
  - id            : "1"
    institution   : "Monash University, Department of Psychology, Jeffrey Cheah School of Medicine and Health Sciences, Malaysia"
  - id            : "2"
    institution   : "Psychiatric Hospital of the University of Basel (UPK), Centre for Chronobiology, Basel, Switzerland"
  - id            : "3"
    institution   : "University of Basel, Transfaculty Research Platform Molecular and Cognitive Neurosciences, Basel, Switzerland"
  - id            : "4"
    institution   : "Eindhoven University of Technology, Department of the Built Environment, Building Lighting, Eindhoven, Netherlands"
  - id            : "5"
    institution   : "Eindhoven University of Technology, Intelligent Lighting Institute, Eindhoven, Netherlands"
  - id            : "6"
    institution   : "Eindhoven University of Technology, Department of Industrial Engineering and Innovation Sciences, Human-Technology Interaction, Eindhoven, Netherlands"
  - id            : "7"
    institution   : "Monash University, Department of Electrical and Computer Systems Engineering, Selangor, Malaysia"
  - id            : "8"
    institution   : "University of Colorado Boulder, Department of Integrative Physiology, Boulder, USA"

  - id            : "9"
    institution   : "XIMES GmbH, Vienna, Austria"
  - id            : "10"
    institution   : "Uniformed Services University of the Health Sciences, Department of Psychiatry, Bethesda, USA"   
  - id            : "11"
    institution   : "Translational Sensory & Circadian Neuroscience, Max Planck Institute for Biological Cybernetics, Tübingen, Germany"
  - id            : "12"
    institution   : "TUM Department of Sport and Health Sciences (TUM SG), Technical University of Munich, Munich, Germany"
  - id            : "13"
    institution   : "University of Oxford, Department of Experimental Psychology, Oxford, United Kingdom"
  - id            : "*"
    institution   : "Joint first author"

authornote: |
  This research is supported by funding from the Welcome Trust (204686/Z/16/Z), the European Training Network LIGHTCAP (project number 860613) under the Marie Skłodowska-Curie actions framework H2020-MSCA-ITN-2019, the BioClock project (number 1292.19.077) of the research program Dutch Research Agenda: Onderzoek op Routes door Consortia (NWA-ORC) which is (partly) financed by the Dutch Research Council (NWO), and the European Union and the nationals contributing in the context of the ECSEL Joint Undertaking programme (2021-2024) under the grant #101007319.

abstract: |
  Light exposure is an important driver of health and well-being. Many aspects of light exposure are modulated by our behaviour. How these light-related behaviours can be shaped to optimise personal light exposure is currently unknown. Here, we present a novel, self-reported and psychometrically validated instrument to capture light exposure-related behaviour, the Light Exposure Behaviour Assessment (LEBA).

  An expert panel prepared the initial 48 item pool spanning different light exposure related behaviours. Responses, consisting rating the frequency of engaging in the per-item behaviour on a 5-point Likert type scale were  collected in an online survey yielding responses from a geographically unconstrained sample (690 completed responses, 74 countries, 28 time zones). Exploratory factor analysis (EFA) on an initial subset of our sample (n=428) rendered a five-factor solution with 25 items (Wearing blue light filters, spending time outdoors, using phone and smart-watch in bed, using light before bedtime, using light in the morning and during daytime). In a confirmatory factor analysis (CFA) performed on an independent subset of participants (n=262), we removed two further items to attain the best fit for the five-factor solution (CFI=0.95, TLI=0.95, RMSEA=0.06). The internal consistency reliability coefficient for the total instrument was, McDonald’s Omega(total)=0.68. Measurement model invariance analysis between native and non-native English speakers showed our model attained the highest level of invariance (residual invariance; CFI=0.95, TLI=0.95, RMSEA=0.05). Lastly, a short form of LEBA (n=18) was developed using Item Response Theory on the complete sample (n=690).

  The psychometric properties of the LEBA instrument indicate the usability to measure the light exposure-related behaviours across a variety of settings and may offer a scalable solution to characterise light exposure-related behaviours in remote samples. The LEBA instrument will be available under the open-access CC-BY-NC-ND license.

keywords          : "light exposure, light-related behaviours, non-visual effects of light, psychometrics"
wordcount         : "X"

bibliography      : ["references.bib", "lib_references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : 
  papaja::apa6_pdf:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: header.tex
header-includes:
  - \usepackage{rotating}
  - \DeclareDelayedFloatFlavor{sidewaysfigure}{figure}

mainfont: Arial
---

```{r setup, warning=FALSE, include=FALSE}
#This chunk controls the common chunk parameters for the whole manuscript
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, dpi=600, fig.width = 6,
 fig.asp = 0.8,out.width = "80%", dev=c('png','postscript'))
#options(knitr.duplicate.label = "allow")
set.seed(123)
par(family = "Arial")
```

```{r pacman, eval=FALSE, include=FALSE}
#This chunk holds code for 'pacman' based library management.

#You need to run this chunk 1X in your system 
#install.packages("pacman")#Run this if you don't have pacman installed.
pacman::p_load(MOTE, tidyverse, psych, lavaan,kableExtra,gt,gtsummary, mirt,likert,kutils,semPlot,semTable,semTools,ggcorrplot,dlookr, paran,EFA.MRFA,VIM,DiagrammeR,DiagrammeRsvg,ggplot2,cowplot,questionr,magick, simsem,readxl, stringr,RColorBrewer,WrightMap,reshape, ggsci,ggtext)
webshot::install_phantomjs()
pacman::p_install_gh("jthomasmock/gtExtras")
pacman::p_install_gh("crsh/papaja")
pacman::p_install_gh("masiraji/tabledown")
pacman::p_install_gh("crsh/citr")
devtools::install_github("unDocUMeantIt/koRpus")
koRpus::install.koRpus.lang(lang=c("en"))

install.packages("correlation")
install.packages("ggtext")
#if you don't have LaTeX installed run the following lines
#install.packages("tinytex")
#tinytex::install_tinytex()
```

```{r library, include=FALSE}
#This chunk holds code for calling the required packages

library(papaja) # devtools::install_github("crsh/papaja")
library(lavaan)
library(semPlot) #devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
library(semTools)
library(MOTE)
library(car)#reverse-coding 
library(psych)
library(dlookr)
library(tidyverse)
library(qgraph)
library(kableExtra)
library(paran) # Parallel analysis
library(EFA.MRFA) # Hull method
library(ggcorrplot)
library(semTable)
library(mirt)
library(gtsummary)
library(gt)
library(gtExtras)
library(likert)
library(VIM) #Missing data
library(kutils)
library(tabledown)
library(readxl)
library(koRpus)# stable release
library(koRpus.lang.en)
library(reshape)
library(ggsci)
library(ggtext)
r_refs("lib_references.bib")
```

```{r ggplot2, include=F}
#This chunk holds code for creating ggplot2 based apatheme for plots.
apatheme=theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        #text=element_text(family = "Helvetian"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        plot.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title=element_blank(),
        axis.line.x = element_line(color='black'),
        axis.line.y = element_line(color='black'),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))

```

---
nocite: |
  `r cite_r("lib_references.bib")`
  [@verriotto2017new][@eklund1996development][@bajaj2011validation][@dianat2013objective]
  [@horne1976self][@roenneberg2003life][@buysse1989pittsburgh][@Xie.2021][@bossini2006sensibilita][@grandner2014development]
---

# Introduction

Light exposure received by the eyes affects many facets of human health, well-being, and performance beyond visual sensation and perception [@boyce_light_2022]. The so-called non-image-forming (NIF) effects of light comprise light’s circadian and non-circadian influence on several physiological and psychological functions, such as the secretion of melatonin, sleep, mood, pupil size, body temperature, alertness, and higher cognitive functions [i.a. reviewed in @Bedrosian.2017;@lok2018light; @siraji2021effects; @blume_effects_2019; @paul_direct_2019; @santhi_applications_2020; @zele_editorial_2020 ].
With the introduction of artificial electric light, human behaviour has become somewhat independent of the natural light-dark cycle – people can now frequently choose when to be exposed to light or darkness. For example, they can decide whether to go outdoors and seek out sunlight, switch on/off light-emitting devices, use certain types of lights at home, or avoid specific light environments altogether. Additionally, when light sources can not be directly manipulated, sought out, or avoided (for example, at school, work, or in public places), there is still potential leeway to influence them behaviourally, for instance, by wearing sunglasses, directing one’s gaze away or supplementing the situation with additional light sources. Although clearly yielding the potential for good, this agency is further associated with increased electric light exposure at night and indoor time during the day, compromising the natural temporal organisation of the light-dark cycle. For example, in the US, an average of 87% of the time is spent in enclosed buildings [@Klepeis.2001], and more than 80% of the population is exposed to a night sky that is brighter than nights with a full moon due to electric light at night [@navara_dark_2007].
An extensive body of scientific evidence suggests that the imbalance of light and dark exposure disrupts humans’ light-dependent physiological systems [@Lunn.2017]. This, in turn, gives rise to a series of adverse health consequences, including the alteration of several hormonal rhythms, increased cancer rates, cardiovascular diseases, and metabolic disorders, such as obesity, and type II diabetes [see reviews by @Lunn.2017; @navara2007dark; @Chellappa.2019]
These findings have sparked a significant call for assessment and guidance regarding healthy light exposure and timing – the latter was recently published as consensus-based experts’ recommendations, postulating specific requirements for indoor light environments during the daytime, evening, and nighttime [@Brown.2022]. Furthermore, building on earlier attempts [e.g. @hubalek_ambulant_2006], there was a recent push toward the development and use of portable light loggers to improve ambulant light assessment and gain more insight into the NIF effects of light on human health in field conditions [@stampfli_light-dosimeter_2021; @van_duijnhoven_recommendations_2017; @webler_towards_2021]. Attached to different body parts (e.g., wrist, head on eye level, chest), these devices allow objectively measuring personal light exposure under real-world conditions and are valuable tools for field studies. Nevertheless, these devices also encompass limiting factors such as potentially being intrusive (e.g., when eye-level worn), yielding the risk of getting covered (e.g., when wrist- or chest-worn) and requiring (monetary) resources and expertise for acquisition and maintenance of the devices.
On the other hand, several attempts have been made to quantify received light exposure subjectively with self-report questionnaires (cf. **Supplementary Table 1**), bypassing the cost and intrusiveness issues. However, subjective light intensity assessments pose a new set of challenges: The human visual system constantly adapts to brightness [@hurvich_perception_1966], while the human non-visual light processing works largely subconsciously [@allen_exploiting_2018], making the self-report assessment of light properties potentially quite challenging, especially for inexperienced laypeople. Retrospectively recalling the properties of a light source can further complicate such subjective evaluations. Moreover, measuring light properties alone does not yield any information about how individuals might behave differently regarding diverse light situations. These measurement limitations point to a couple of research challenges we aim to take on here: How can we gain insight into light exposure patterns via self-report but circumvent directly inquiring about the specific properties and intensity of a light source? And how can we simultaneously assess how people habitually interact with the received light?
We propose that these challenges can be tackled by assessing light-exposure-related behaviour. We argue that, besides measuring received light exposure as intensity, it is also essential to understand people's behaviours concerning different light situations. Since, in many cases, humans have become their own agents regarding their exposure to light or darkness through artificial electric light, people's light exposure-related behaviours ultimately determine their light consumption and timing: People receive different light depending on their daily activities, including workplace habits, bedtime hygiene, pastime and social activities. The final objective of changing light-dark exposure patterns to avoid or mitigate negative health consequences from unhealthy habits will not just need an assessment of the lighting properties but the active change of behaviours related to light exposure. We argue that assessing these activities is a beneficial stepping stone for prospective behaviour change. Furthermore, people without light measurement expertise may find it easier to appraise and recall their behaviour concerning light exposure than subjectively assessing a light source’s properties.
To date, little effort has been made to understand and capture these activities. **Supplementary Table 1** summarises the existing questionnaire literature assessing light exposure-related properties. However, only a few questions of these existing tools were associated with light exposure-related behaviour. For example, the “Munich Chronotype  Questionnaire” [MCTQ; @roenneberg2003life], a popular self-report tool for identifying chronotypes via mid-sleep times, includes questions about the individual’s time spent outdoors on workdays and free days. In addition, the “Pittsburgh Sleep Quality Index” [PSQI; @buysse1989pittsburgh], a popular measure of sleep quality, contains questions about sleep and wake-up times, which are relevant to the light exposure around bedtime. However, none of these questionnaires provides a scaleable solution to capture light exposure-related behaviour in various physiologically relevant lighting scenarios.
To fill this gap, we here present the development process of a novel self-report tool - the “Light Exposure Behavior Assessment” (LEBA) - for capturing and quantifying diverse light exposure-related behaviours.

# Methods

```{r Data, include=FALSE}
#This chunk holds code for data wrangling

data <- readRDS("leba_2021-09-08.rds")

# Separating EFA and CFA samples with descriptive column
descriptives.data <- data

## Merge "0"s and "1"s into "1"s select,subset assign)
descriptives.data[ , 9:56 ][ descriptives.data[ , 9:56 ] == 0 ] <- 1
#EFA.descriptives <- descriptives.data[1:428,]
EFA.descriptives <- subset(descriptives.data, IncludedInEFA == "TRUE")
#CFA.descriptives <- descriptives.data[429:690, ]
CFA.descriptives<- subset(descriptives.data, IncludedInEFA == "FALSE")

#Separating the EFA and CFA(only items)
sem.data <- data[, 9:56] #EFA & CFA data
sem.data[ sem.data == 0] <- 1 #Merged "0"s and "1"s into "1"s

invariance.data.descriptives.count <- CFA.descriptives %>% 
  group_by(slypos_demographics_language.factor) %>% 
  count() %>% 
  as.data.frame()

## renaming the column-header
prefix <- "item"
sufix <- c(1:48)
colnam <- paste(prefix,sufix, sep = "" )
colnames(sem.data) <- colnam
names(sem.data)[names(sem.data) == "item1"] <- "item01"
names(sem.data)[names(sem.data) == "item2"] <- "item02"
names(sem.data)[names(sem.data) == "item3"] <- "item03"
names(sem.data)[names(sem.data) == "item4"] <- "item04"
names(sem.data)[names(sem.data) == "item5"] <- "item05"
names(sem.data)[names(sem.data) == "item6"] <- "item06"
names(sem.data)[names(sem.data) == "item7"] <- "item07"
names(sem.data)[names(sem.data) == "item8"] <- "item08"
names(sem.data)[names(sem.data) == "item9"] <- "item09"
```

```{r country, include=FALSE}
#This chunk holds code for time-zone calculation

library(stringr)
country <- data

TZ <- as.data.frame(str_split_fixed(country$slypos_demographics_tz.factor, "-", 2))
TZ[97, 2] <- " East Africa/Dodoma (UTC +03:00)"
TZ[146, 2] <- " European /Skopje (UTC +01:00)"
TZ[176, 2] <- " Asia /Taipei City (UTC +08:00)"
TZ[574, 2] <- " Asia /Taipei City (UTC +08:00)"
TZ[690, 2] <- " Asia /Taipei City (UTC +08:00)"
TZ[273, 2] <- " Iran /Tehran (UTC +0:30)"
TZ[418, 2] <- " Iran /Tehran (UTC +0:30)"
TZ[673, 2] <- " Iran /Tehran (UTC +0:30)"

num.countries <- as.data.frame(unique(TZ$V1))
num.Timezone <- unique(TZ$V2)

UTC <- as.data.frame(str_split_fixed(TZ$V2, "UTC", 2))
numUTC <- as.data.frame(unique(UTC$V2))
nrow(numUTC)
```

```{r timezone, include=F}
#This chunk holds code for creating time-zone.csv file
timezone <- TZ %>% 
  group_by(V2) %>% 
  count() %>% 
  as.data.frame()

write.csv(timezone, "Table_raw/timezone.csv")

```

```{r prepareDescTable, include=FALSE}
#This chunk holds code for demographic data wrangling.

listdescVars <- colnames(descriptives.data[c(1:3,5:7)])

#separating naming and  reducing data for descriptive Table 
desctable.data <- dplyr::select(descriptives.data, c(listdescVars,
                          IncludedInCFA))

#renaming TRUE and FALSE from the IncludedInCFA to EFA and CFA for Descriptive Table
desctable.data$IncludedInCFA[desctable.data$IncludedInCFA == T ] <-
 "2. CFA Sample"
desctable.data$IncludedInCFA[desctable.data$IncludedInCFA == F ] <- 
 "1. EFA Sample"

# recode the Gender factor so it will only show "Gender Diverse" in the summary table
desctable.data$slypos_demographics_gender.factor <-
  recode_factor(desctable.data$slypos_demographics_gender.factor, "No" = "Yes",
         "Yes"= "No")

# create descriptive table for the demographic vars (excluding tz & Country) with gtsummary
desctable.data %>%
  tbl_summary(
    by = IncludedInCFA,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    digits = all_continuous() ~ 2,
    label = list(slypos_demographics_age ~ "Age",
                 slypos_demographics_sex.factor ~ "Sex",
                 slypos_demographics_gender.factor ~ "Gender-Variant Identity",
                 slypos_demographics_language.factor ~ "Native English Speaker",
                 slypos_demographics_work_or_school.factor ~ "Occupational Status",
                 slypos_demographics_school.factor ~ "Occupational setting"
                 ),

    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% 
  # add_p() %>% add_q() %>%
modify_header(label ~ "**Variable**") %>% 
  # separate_p_footnotes() %>%
  modify_caption("Demographic Characteristics of Participants (n=690).") -> desc_table

as_tibble(desc_table) -> desc_tibble 
as_kable_extra(desc_table, format = "latex",booktabs = T) -> desc_kable #save it as a knitr::kable

#summarise country/time zone data
#The following chunk is not working now. May be package problem (Sorting command is not working)

descriptives.data%>%
  tbl_summary(
    label = slypos_demographics_tz.factor ~ "Time zone - Country",
    statistic = list(all_categorical() ~ "{n} ({p}%)"),
    missing = "no",
    sort = all_categorical() ~ "frequency",
    include = "slypos_demographics_tz.factor"
  )  %>% bold_labels() %>%
  modify_header(label ~ "") %>%
  as_tibble(format = 'pipe') -> tz_tibble

```

## Data Collection

```{r demotab, warning=F, message=F}
desc_kable  %>% kable_styling(latex_options = c("scale_down")) %>% landscape()
```

A quantitative cross-sectional fully anonymous, geographically unconstrained online survey was conducted via REDCap [@harris2009research; @harris2019redcap] by way of the University of Basel [sciCORE](https://redcap.scicore.unibas.ch). Participants were recruited via the [website](https://enlightenyourclock.org/participate-in-research) (<https://enlightenyourclock.org/participate-in-research>) of the science-communication comic book "Enlighten your clock", co-released with the survey [@Weinzaepflen.2021], social media (i.e., LinkedIn, Twitter, Facebook), mailing lists, word of mouth, the investigators' personal contacts, and supported by the distribution of the survey link via f.lux [@f.lux]. The initial page of the online survey provided information about the study, including that participation was voluntary and that respondents could withdraw from participation at any time without being penalised. Subsequently, consent was recorded digitally for the adult participants (\>18 years), while under-aged participants (\<18 years) were prompted to obtain additional assent from their parents/legal guardians. Filling in all questionnaires was estimated to take less than 30 minutes, and participation was not compensated.
As a part of the demographic data, participants provided information regarding age, sex, gender identity, occupational status, COVID-19-related occupational setting, time zone/country of residence and native language. The demographic characteristics of our sample are given in **Table 1**. Participants were asked to confirm that they participated in the survey for the first time. Additionally, five attention check items (e.g., "We want to make sure you are paying attention. What is 4+5?") were included among the questionnaires to ensure high data quality. All questions incorporating retrospective recall were aligned to a "past four weeks" period.

We collected the survey data between 17 May 2021 and 3 September 2021 – firstly from `r nrow(EFA.descriptives)` participants (EFA sample) – and subsequently, another dataset from `r nrow(CFA.descriptives)` participants (CFA sample), totalling `r nrow(desctable.data)`.

## Analytic Strategy

Figure \@ref(fig:FlowchartFig) summarises the steps we followed while developing LEBA. We conducted all analyses with the statistical software environment R [@R-base]. 
**Firstly**, we set an item pool of 48 items with a six-point Likert-type response format (0-Does not apply/I don't know, 1-Never, 2-Rarely 3-Sometimes, 4-Often, 5-Always) for our initial scale. Our purpose was to capture light exposure-related behaviour. In that context, the first two response options: "Does not apply/I don't know" and "Never" provided similar information. As such we collapsed them into one, making it a 5-point Likert-type response format (1-Never, 2-Rarely, 3-Sometimes, 4-Often, 5-Always).

**Secondly**, the two rounds of data collection were administered. 
**Thirdly**, we conducted descriptive and item analysis and proceeded to the exploratory factor analysis (EFA) using the "psych" package [@R-psych] on the data collected in the first round (EFA sample; n=`r nrow(EFA.descriptives)`), as a part of psychometric analysis. Prior to the EFA, the necessary assumptions, including sample adequacy, normality assumptions, and quality of correlation matrix, were assessed. As our data violated both the univariate and multivariate normality assumption and yielded ordinal response data, we used a polychoric correlation matrix in the EFA and employed "principal axis" (PA) as the factor extraction method [@desjardinsHandbookEducationalMeasurement2018; @watkinsStepbyStepGuideExploratory2020]. We applied a combination of methods including Scree plot [@cattellScreeTestNumber1966], minimum average partials method [@velicerDeterminingNumberComponents1976], and Hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. To determine the latent structure, we followed the common guidelines: (i) no factors with fewer than three items (ii) no factors with a factor loading \<0.3 (iii) no items with cross-loading \> .3 across factors [@bandalosFactorAnalysisExploratory2018]. <!-- We also conducted an EFA on non-merged response options data set (**Supplementary File 1**). -->

For reliability estimation, the "psych" package was applied [@R-psych]. Though Cronbach's internal consistency coefficient alpha is widely used for estimating internal consistency, it tends to deflate the estimates for Likert-type data since the calculation is based on the Pearson-correlation matrix, which requires response data to be continuous in nature [@gadermann2012estimating; @zumbo2007ordinal]. Subsequently, we reported ordinal alpha for each factor obtained in the EFA [@zumbo2007ordinal] to get better reliability estimates. We also estimated the internal consistency reliability of the total scale using McDonald's $\omega_t$ coefficient, which was suggested as a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. Both ordinal alpha and McDonald's $\omega_t$ coefficient values range between 0 to 1, where higher values represent better reliability.

To validate the latent structure obtained in the EFA, we conducted a categorical confirmatory factor analysis (CFA) with the weighted least squares means and variance adjusted (WLSMV) estimation [@desjardinsHandbookEducationalMeasurement2018], using the "lavaan" package [@R-lavaan] on the data collected in the second round (CFA sample; n=`r nrow(CFA.descriptives)`). We assessed the model fit using common model fit guidelines: (i) $\chi^2$ test statistics: a non-significant test statistics is required to accept the model (ii) comparative fit index (CFI) and Tucker Lewis index (TLI): close to .95 or above/ between .90-.95 and above (iii) root mean square error of approximation (RMSEA): close to .06 or below, (iv) Standardized root mean square (SRMR): close to .08 or below [@huCutoffCriteriaFit1999; @schumacker2004beginner]. However, the $\chi^2$ test is sensitive to sample size [@brownConfirmatoryFactorAnalysis2015], and SRMR does not work well with ordinal data [@RN1272]. Consequently, we judged the model fit using CFI, TLI and RMSEA.

We then assessed the measurement invariance (MI) of our scale between native English speakers (n=`r invariance.data.descriptives.count[1,2]`) and non-native English speakers (n=`r invariance.data.descriptives.count[2,2]`) in the CFA sample (n=`r nrow(CFA.descriptives)`). MI evaluates whether a construct has the psychometric equivalence and the same meaning across groups [@putnick2016measurement; @klinePrinciplesPracticeStructural2015]. We used the structural equation modelling framework applying the "lavaan" package [@R-lavaan] to assess the measurement invariance. We successively compared four nested models: configural, metric, scalar, and residual models using the $\chi^2$ difference test ($\Delta \chi^2$). Among MI models, the configural model is the least restrictive and the residual model is the most restrictive. A non-significant $\Delta \chi^2$ test between two nested measurement invariance models indicates mode fit does not significantly decrease for the superior model, thus allowing the superior invariance model to be accepted [@dimitrov2010testing; @widaman1997exploring].

**Fourthly**, as secondary analysis, we identified the educational grade level required to understand the items in our scale with the Flesch-Kincaid grade level identification method [@flesch1948new] applying the "koRpus" [@R-koRpus] package. Also, we analysed possible semantic overlap of our developed scale using the ["Semantic Scale Network" (SSN)](https://rosenbusch.shinyapps.io/semantic_net/) engine [@rosenbusch2020semantic]. The SSN detects semantically related scales and provides a cosine similarity index ranging between -.66 to 1 [@rosenbusch2020semantic]. Pairs of scales with a cosine similarity index value of 1 inidicate full semantical similarity, suggesting redundancy of the scales.

**Lastly**, we derived a short form of the LEBA employing an Item Response Theory (IRT) based analysis. We fitted each factor of the LEBA to the combined EFA and CFA sample (n=`r nrow(sem.data)`) using the graded response model [@samejima1997handbook] via the "mirt" package [@R-mirt]. IRT assesses the item quality by estimating the item discrimination, item difficulty, item information curve, and test information curve [@bakerBasicsItemResponse2017]. Item discrimination indicates how well a particular item can differentiate between participants across the given latent trait continuum ($\theta$). Item difficulty corresponds to the latent trait level at which the probability of endorsing a particular response option is 50%. The item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported the item difficulty and discrimination parameter and categorize the items based on their item discrimination index: none = 0; very low = 0.01 to 0.34; low = 0.35 to 0.64; moderate = 0.65 to 1.34 ; high = 1.35 to 1.69; very high \>1.70 [@bakerBasicsItemResponse2017]. We discarded the items with a relatively flat item information curve (information \<.2) to derive the short form of LEBA. We also assessed the precision of the short LEBA utilizing the Test information curve (TIC). TIC indicates the amount of information a particular scale carries along the latent trait continuum. Additionally, the item and person fit of the fitted IRT models were analysed to gather more evidence on the validity and meaningfulness of our scale [@desjardinsHandbookEducationalMeasurement2018]. The item fit was evaluated using the RMSEA value obtained from Signed-$\chi^2$ index implementation, where an RMSEA value $\le$.06 was considered adequate item fit. The person fit was estimated employing the standardized fit index Zh statistics [@drasgow1985appropriateness]. Here, Zh \< -2 was considered as a misfit [@drasgow1985appropriateness].

```{r FlowchartFig, echo=FALSE, fig.align='center', fig.cap='Flow chart of the LEBA (long and short form) development and evaluation. ', out.height='100%', out.width='100%'}
knitr::include_graphics('Figures/Figure1.png', dpi = 600 )
```

## Ethical Approval

The current research project utilizes fully anonymous online survey data and therefore does not fall under the scope of the Human Research Act, making an authorisation from the ethics committee redundant. Nevertheless, the cantonal ethics commission (Ethikkommission Nordwest- und Zentralschweiz, EKNZ) reviewed our proposition (project ID Req-2021-00488) and issued an official clarification of responsibility.

## Data Availability

The present article is a fully reproducible open access "R Markdown" document. All code and data underlying this article -- along with two versions of the LEBA questionnaire (full and short) and online survey implementation templates on common survey platforms -- will be available under open-access licence (CC-BY-NC-ND) on a public [GitHub repository](https://github.com/leba-instrument).

# Results

## Development of the Initial Scale

An expert panel comprising all authors -- researchers from chronobiology, light research, neuroscience and psychology -- developed a comprehensive item pool of 48 items. The 48 items were then examined based on their relevance and representativeness of the construct "Light Exposure Related Behaviour" by each panel member independently and modifications were suggested as required. The author team discussed the suggestions and amended the items as indicated, thus creating a 48-item scale.

## Anonymous Online Survey

Table 1 summarises the survey participants' demographic characteristics. Only participants completing the full LEBA questionnaire were included. Thus, there are no missing values in the item analyses. (XXX??) participants were excluded from the analysis due to not passing at least one of the "attention check" items. For the EFA, a sample of at least 250-300 is recommended [@comreyFirstCourseFactor1992; @schonbrodtWhatSampleSize2013]. To assess sampling adequacy for CFA, we followed the N:q rule [@bentlerPracticalIssuesStructural1987; @jacksonRevisitingSampleSize2003; @klinePrinciplesPracticeStructural2015; @worthingtonScaleDevelopmentResearch2006], where at least ten participants per item are required to earn trustworthiness of the result. Both our EFA and CFA sample size exceeded these requirements. Participants indicated filling out the online survey from various geographic locations, including `r nrow(num.countries)` countries and `r nrow(numUTC)` time zones. For a complete list of geographic locations, see **Supplementary Table 2**.

Participants in our survey were aged between `r min(desctable.data$slypos_demographics_age)` to `r max(desctable.data$slypos_demographics_age)` years, with an overall mean of \~ `r round(mean(desctable.data$slypos_demographics_age), 2)` years of age [Overall: `r mean(desctable.data$slypos_demographics_age)`±`r sd(desctable.data$slypos_demographics_age)`; EFA: `r mean(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`±`r sd(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "1. EFA Sample"])`; CFA: `r mean(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`±`r sd(desctable.data$slypos_demographics_age[desctable.data$IncludedInCFA == "2. CFA Sample"])`]. In total `r desc_tibble[3,2]` of the participants indicated female sex, `r desc_tibble[4,2]` indicated male and `r desc_tibble[5,2]` indicated other sex. Overall, `r desc_tibble[6,2]` participants indicated a gender-variant identity. In a "Yes/No" question regarding native language, `r desc_tibble[7,2]` of respondents [EFA: `r desc_tibble[7,3]`; CFA: `r desc_tibble[7,4]`] indicated to be native English speakers. For their "Occupational Status", more than half of the overall sample reported that they currently work, whereas `r desc_tibble[10,2]` reported that they go to school and `r desc_tibble[11,2]` responded that they do "Neither". With respect to the COVID-19 pandemic we asked participants to indicate their occupational setting during the last four weeks: In the overall sample `r desc_tibble[13,2]` of the participants indicated that they were in a home office/ home schooling setting, while `r desc_tibble[14,2]` reported face-to-face work/schooling. Lastly, `r desc_tibble[15,2]` overall reported a combination of home- and face-to-face work/schooling, whereas `r desc_tibble[16,2]`  filled in the "Neither (no work or school, or on vacation)" response option.

## Psychometric Analysis: Development of the Long Form

### Descriptive Statistics and Item Analysis

```{r full-data-mardia, include =F}
mardia.all.data <- psych::mardia(sem.data, na.rm = T, plot =T)
descriptives.all.data <- tabledown::des.tab(sem.data)
```

```{r gtvis, include =F}
#This chunk holds codes for creating 'gtExtra' based descriptive figures (data preparation)

## Recoding sem.data$item8
gt.data <- sem.data
gt.data$item08 <- as.character(gt.data$item08)
gt.data$item08 <- fct_recode(gt.data$item08,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)

## Recoding gt.data$item26
gt.data$item26 <- as.character(gt.data$item26)
gt.data$item26 <- fct_recode(gt.data$item26,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)
## Recoding gt.data$item37
gt.data$item37 <- as.character(gt.data$item26)
gt.data$item37 <- fct_recode(gt.data$item26,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)

# all data Gt Table
##Long table
gt.long <- as.data.frame(gather(gt.data, Items, value))
gt.long$value <- as.numeric(as.character(gt.long$value))

##Summarizing and creating gt object

gt.tab <- gt.long %>% 
  group_by(Items) %>% 
  # calculate summary stats & create data for the histogram and density plot
  dplyr::summarise(
    nr = n(),
    mean = mean(value, na.rm = TRUE),
    # med = median(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    hist_data = list(value),
    dens_data = list(value),
    .groups = "drop"
  ) %>% 
  gt()

gt.tab1 <- gt.tab$`_data`

gt.tab2 <- gt.tab1[,-c(1,2)]

vars_labels = as.data.frame(sapply(sem.data,
       function(x){attr(x,"label")}))

vars_labels<- tibble::rownames_to_column(vars_labels, "Items")
colnames(vars_labels) <- c("Item", "Stem")

gt.tab3 <- cbind(vars_labels, gt.tab2[,c(1,2)], descriptives.all.data$Normality,gt.tab2[,c(3,4)] )

colnames(gt.tab3) <- c("Item", "Stem", "mean", "sd","S-W Statistics", "hist_data","dens_data"  )
#Preparation for likert data
LEBA.likert <- as.data.frame(gt.data)

recod_LEBA <- c( "1" = "Never", "2" = "Rarely", "3"= "Sometimes","4" = "Often",
                 "5" = "Always")
LEBA.likert <-  mutate(LEBA.likert, across(starts_with("item"), ~unname(recod_LEBA[.])))

LEBA.Factor = as.data.frame(lapply(LEBA.likert,factor,
                                   ordered = T))
#get the items name
items <- names(LEBA.Factor) 
#Calculate percentage
percentage <- kutils::likert(LEBA.Factor, vlist = items )

percentage <- percentage$table %>% 
  as.data.frame(.)

#data wrangling
labels <- c("Always", "Never","Often", "Rarely", "Sometimes","Total")
as.data.frame(labels)

full.percentage <- cbind(labels,percentage) #tables with labels  
full.percentage<- t(full.percentage ) #transpose
as.data.frame(full.percentage)
full.percentage1 <- full.percentage[-1,-6] #removing 1st row and total column
full.percentage2 <- full.percentage1[, c(2, 4, 5, 3,1)]# rearranging
as.data.frame(full.percentage2)

colnames(full.percentage2) <- c("Never","Rarely","Sometimes","Often","Always")

Items <- rownames(full.percentage2)
as.data.frame(Items)
full.percentage3 <- cbind(Items,full.percentage2)

full.percentage3 <- full.percentage3[order(Items),] 
full.percentage3 <- as.data.frame(full.percentage3[,-1]) %>% 
  gt()

liket.full <- full.percentage3$"_data"

full.table <-  cbind( gt.tab3, liket.full)%>% 
  gt()

```

```{r fullgt,  include=FALSE}
#This chunk holds codes for creating 'gtExtra' based descriptive figures (EFA+CFA);

full.tab <- full.table$`_data`

full.tab.1 <- full.tab[1:24,] %>% 
  gt()

full.tab.2 <- full.tab[25:48,] %>% 
  gt()

full.tab.1 %>% 
# histogram and density plots
  gtExtras::gt_sparkline(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_sparkline(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "grey",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("Summary Descriptives (n=690)"),
    subtitle = md("Items 01-24 ")) %>% 
#create groups of columns
tab_spanner(
  label = "Item",
  columns = Item:Stem
) %>%
  cols_align(
    align = "left",
    columns = Item:Stem
  ) %>% 
  tab_spanner(
    label = "Summary Statistics",
    columns = mean:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%

     tab_footnote(
     footnote = md("**Shapiro–Wilk test**"),
     locations = cells_column_labels(columns = `S-W Statistics`)
  ) %>% 
# change column names to appear in the table
cols_label(
  Item = ("Items"),
  Stem = ("Stem"),
  mean = ("Mean"),
  sd = (("SD")),
  `S-W Statistics` = ("SW"),
  hist_data = "Histogram",
  dens_data = "Density"
)  %>% 
# set alignment as per wish
  cols_align(
    align = "center",
    columns = mean:Always
  ) %>%
  opt_align_table_header(align = "left") %>%
# add coloured dots and lines on the first column
  gt_plt_dot(
    mean,
    Item,
    palette = "ggthemes::fivethirtyeight"
  ) %>% tab_options(
    table.font.size = px(14L)) %>% 
  cols_width(
    Stem ~ px(200)) %>% 
 gtsave("Figures/Figure2.png", vwidth = 6000)

full.tab.2 %>% 
# histogram and density plots
  gtExtras::gt_sparkline(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_sparkline(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "grey",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("Summary Descriptives (n=690)"),
    subtitle = md("Items 25-48 ")) %>% 
#create groups of columns
tab_spanner(
  label = "Item",
  columns = Item:Stem
) %>%
  cols_align(
    align = "left",
    columns = Item:Stem
  ) %>% 
  tab_spanner(
    label = "Summary Statistics",
    columns = mean:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%

     tab_footnote(
     footnote = md("**Shapiro–Wilk test**"),
     locations = cells_column_labels(columns = `S-W Statistics`)
  ) %>% 
# change column names to appear in the table
cols_label(
  Item = ("LEBA  Items"),
  Stem = ("Stem"),
  mean = ("Mean"),
  sd = (("SD")),
  `S-W Statistics` = ("SW"),
  hist_data = "Histogram",
  dens_data = "Density"
)  %>% 
# set alignment as per wish
  cols_align(
    align = "center",
    columns = mean:Always
  ) %>%
  opt_align_table_header(align = "left") %>%
# add coloured dots and lines on the first column
  gt_plt_dot(
    mean,
    Item,
    palette = "ggthemes::fivethirtyeight"
  ) %>% tab_options(
    table.font.size = px(14L)) %>% 
  cols_width(
    Stem ~ px(200)) %>% 
gtsave("Figures/Figure3.png",vwidth = 6000)

```

```{r efagtPic1, echo=FALSE,  fig.cap= 'Summary descriptives and response pattern observed in the large-scale survey for item 01-24. All items violated normality assumption.', out.height='100%', out.width='250%'}
knitr::include_graphics('Figures/Figure2.png')

```

```{r efagtPic2, echo=FALSE,  fig.cap= 'Summary descriptives and response pattern observed in the large-scale survey for item 25-48. All items violated normality assumption.', out.height='100%', out.width='250%'}
knitr::include_graphics('Figures/Figure3.png')

```

Figure \@ref(fig:efagtPic1) and Figure \@ref(fig:efagtPic2) summarise the response patterns of our total sample (n=`r nrow(sem.data)`) for all 48 items. Most of the items appeared skewed. The Shapiro–Wilk test of univariate normality [@shapiroAnalysisVarianceTest1965] and Mardia test of multivariate normality [@mardiaMeasuresMultivariateSkewness1970] indicated our data violated both univariate and multivariate normality. The multivariate skew was `r printnum(mardia.all.data$b1p)` (p \<0.001), and the multivariate kurtosis was `r printnum(mardia.all.data$b2p)` (p \<0.001).

```{r EFAdata, include=FALSE}
#This chunk holds code for creating EFA data subset
### EFA data
EFA.data <- sem.data[1:428, ]
#library(VIM)
missing <- VIM::aggr(EFA.data, plot =T)

```

```{r EFAassumptions, include=FALSE}
#This chunk holds code for checking assumptions of EFA

#KMO test
KMO <- psych::KMO(EFA.data)

# Test of correlation matrix
bartlet <- psych::cortest.bartlett(EFA.data, n =428)

#Histogram
psych::multi.hist(EFA.data[,sapply(EFA.data, is.numeric)])

# Univariate normality

descriptives <- tabledown::des.tab(EFA.data)

colnames(descriptives) <- c("Items", "Mean", "SD", "Skew", "Kurtosis", "SW", "Item Total Correlation")

vars_labels = as.data.frame(sapply(EFA.data,
       function(x){attr(x,"label")}))

vars_labels<- tibble::rownames_to_column(vars_labels, "Items")
colnames(vars_labels) <- c("Item", "Stem")

descriptives2 <- cbind(vars_labels, descriptives[,-1])
# Multivariate Normality
mardia <- psych::mardia(EFA.data, na.rm = T, plot =T)

```

```{r ItemAnalysis, include=FALSE}
#This chunk holds code for Item analysis (Classical Test Theory)
# Item analysis
Item_analysis <- psych::alpha(EFA.data,check.keys=TRUE)
Item_analysis$item.stats$r.cor
low.r.corec <- (min(Item_analysis$item.stats$r.cor))  #minimum item total correlation
high.r.corec <- (max(Item_analysis$item.stats$r.cor))     # maximum item total correlation
```

```{r gt-EFA, include =F}
#EFA data
gt.efa.data <- EFA.data

# gt.efa.data.table <- gt.efa.data %>% 
#   dplyr::select(item16, item36, item17,
#                  item11, item10,item12,item07,item08,item09,
#                  item27,  item03, item40,  item30, item41,
#                  item33, item32,item35, item37,item38, 
#                  item46, item45, item25, item04, item01,item26)

#Creating long data
gt.efa.long <- as.data.frame(gather(gt.efa.data, Items, value))
gt.efa.long$value <- as.numeric(as.character(gt.efa.long$value))

#Summarizing and creating gt object
gt.efa.tab <- gt.efa.long %>% 
  group_by(Items) %>% 
  # calculate summary stats & create data for the histogram and density plot
  dplyr::summarise(
    # nr = n(),
    # mean = mean(value, na.rm = TRUE),
    # med = median(value, na.rm = TRUE),
    # sd = sd(value, na.rm = TRUE),
    hist_data = list(value),
    dens_data = list(value),
    .groups = "drop"
  ) %>% 
  gt()

#Preparation for likert data
efa.LEBA.likert <- as.data.frame(gt.efa.data)

recod_LEBA <- c( "1" = "Never", "2" = "Rarely", "3"= "Sometimes","4" = "Often",
                 "5" = "Always")
efa.LEBA.likert <-  mutate(efa.LEBA.likert, across(starts_with("item"), ~unname(recod_LEBA[.])))

efa.LEBA.Factor = as.data.frame(lapply(efa.LEBA.likert,factor,
                                   ordered = T))
#get the items name
efa.items <- names(efa.LEBA.Factor) 
#Calculate percentage
efa.percentage <- kutils::likert(efa.LEBA.Factor, vlist = efa.items )

efa.percentage <- efa.percentage$table %>% 
  as.data.frame(.)

#data wrangling
labels <- c("Always", "Never","Often", "Rarely", "Sometimes","Total")
as.data.frame(labels)

full.efa.percentage <- cbind(labels,efa.percentage) #tables with labels  
full.efa.percentage<- t(full.efa.percentage ) #transpose
as.data.frame(full.efa.percentage)
full.efa.percentage1 <- full.efa.percentage[-1,-6] #removing 1st row and total column
full.efa.percentage2 <- full.efa.percentage1[, c(2, 4, 5, 3,1)]# rearranging
as.data.frame(full.efa.percentage2)

colnames(full.efa.percentage2) <- c("Never","Rarely","Sometimes","Often","Always")

efa.Items <- rownames(full.efa.percentage2)
as.data.frame(efa.Items)
full.efa.percentage3 <- cbind(efa.Items,full.efa.percentage2)

full.efa.percentage3 <- full.efa.percentage3[order(efa.Items),] 
full.efa.percentage3 <- as.data.frame(full.efa.percentage3[,-1]) %>% 
  gt()

gt.efa.table <- gt.efa.tab$'_data'
gt.liket.efa <- full.efa.percentage3$"_data"

efaTable <-  cbind( gt.efa.table, gt.liket.efa)%>% 
  gt()

 efa.gt <- efaTable$"_data"

 efa.gt.2 <- cbind(descriptives2[,-2], efa.gt[,-1])
```

```{r efagt,include=F}

#This chunk holds codes for creating 'gtExtra' based descriptive figures (EFA descriptives Supplementary

efa.gt.2 %>% 
gt() %>% 
# histogram and density plots
  gtExtras::gt_sparkline(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_sparkline(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "#E64B3599",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# # format decimals
#   fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("**Light Exposure Behavior Assessment**"),
    subtitle = md("Summary Descriptives EFA Sample (n=428)")
  ) %>% 
   tab_footnote(
     footnote = md("**Shapiro–Wilk test**"),
     locations = cells_column_labels(columns = SW)
  ) %>%

#create groups of columns
tab_spanner(
  label = "Items",
  columns = Item
)  %>% 
  tab_spanner(
    label = "Summary Statistics",
    columns = Mean:SW
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%
# change column names to appear in the table
cols_label(
  hist_data = "Histogram",
  dens_data = "Density"
) %>%
# set alignment as per wish
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_align_table_header(align = "left") %>%
# add coloured dots and lines on the first column
  gt_plt_dot(
    Mean,
    Item,
    palette = "ggthemes::fivethirtyeight") %>% 
  tab_options(
    table.font.size = px(14L)) %>%

  gtsave("Figures/S1_Fig.png", vwidth=6000)

# tab_row_group(
  #     label = md("**F5: Using light in the morning and during daytime**"),
  #     rows = c(1,3,12,13,24,25) 
  #       ) %>% 
  # tab_row_group(
  #     label = md("**F4: Using light before bedtime**"),
  #     rows = c(16,17,18,20,21) 
  #       ) %>% 
  # tab_row_group(
  #     label = md("**F3: Using phone and smart-watch in bed**"),
  #     rows = c(2,14,15,22,23) 
  #       ) %>% 
  # tab_row_group(
  #     label = md("**F2: Spending time outdoors**"),
  #     rows = c(4,5,6,7,8,9) 
  #       ) %>% 
  # tab_row_group(
  #     label = md("**F1: Wearing blue light filters**"),
  #     rows = c(10,11,19) )

```

**Supplementary Figure 1** summarises the univariate descriptive statistics for the 48 items in the EFA sample (n=`r nrow(EFA.data)`). Likewise, our data violated the univariate [@shapiroAnalysisVarianceTest1965] and multivariate normality assumptions [@mardiaMeasuresMultivariateSkewness1970]. The multivariate skew was `r printnum(mardia$b1p)` (p \<0.001) and the multivariate kurtosis yielded a value of `r printnum(mardia$b2p)` (p \<0.001). The corrected item-total correlation ranged between `r apa(low.r.corec,2,F)` -`r apa(high.r.corec,2,F)`. However, no item was discarded based on descriptive statistics or item analysis.

### Exploratory Factor Analysis and Reliability Analysis

```{r CorrMatrix, include=FALSE}
#This chunk holds code for producing correlation matrix and its details.

#Polychoric Correlation matrix

correlations <- psych::polychoric(EFA.data, correct = 0)

upper <- correlations$rho[upper.tri(correlations$rho, diag = F)]

min.cor <- apa(min((upper)),2,F) #minimum cor.coefficient of the matrix
max.cor <- apa(max((upper)),2,F) # max. correlation coefficient of the matrix

determinets <- det(correlations$rho) 
# Calculating the percentage of correlations higher than .30

BigR = sum(correlations$rho >= abs(.30) & correlations$rho < abs(1.0), na.rm =T)/2 
totR = length(EFA.data)*(length(EFA.data)-1)/2
cor.per <- print (BigR/totR)*100
```

We checked the sampling adequacy by applying Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy on the EFA sample (n=`r nrow(EFA.data)`) [@kaiserIndexFactorialSimplicity1974]. The overall KMO value for 48 items was `r printnum(KMO$MSA)`, which exceeded the cut-off value (.50), indicating an adequate sample size [@hutchesonMultivariateSocialScientist1999]. Additionally, Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`)=`r apa(bartlet$chisq, 2, T)`, p \< .001 implied that the correlations between items were adequate for conducting the EFA. However, only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than \|.30\|., and the inter-item correlation coefficients ranged between `r min.cor` to `r max.cor`. Figure \@ref(fig:efa-plot-print)-A depicts the respective correlation matrix.

```{r tabDes, eval=FALSE, include=FALSE, results='asis'}
descriptives2 %>% 
apa_table(landscape =T,caption ="Univariate Descriptive Statistics for the 48 Items. All Items Violated Univariate Normality Assumption Assessed by Shapiro–Wilk (SW) Test. The  Item-total Correlation Ranges Between .03-.48", escape =F, font_size = "footnotesize", note = ("$^{*}$p<.001, $^{1}$Shapiro–Wilk Statistics"), align = c("p{1cm}","p{10cm}", rep("p{1cm}", ncol(descriptives2)))) 
```

```{r correlation-hits, eval=FALSE, include=FALSE}
#This chunk holds code for plotting the correlation matrix in histogram format

cor.hits <- as.data.frame(melt(upper))
colnames(cor.hits) <- c("value")
corr.histogram <- ggplot(cor.hits,aes(x=value)) + 
  geom_histogram(binwidth=.1, col=I("black"),fill="#00A08799")+apatheme+
  labs( y="", x = "Inter-item correlation")

# ggsave("Figures/correlation_histogram.png", corr.histogram, width = 6, height = 6, dpi = 600, bg ="white")
```

```{r corplot, include =F}
#This chunk holds code for plotting the correlation matrix

p.mat <- correlation::cor_to_p(correlations$rho, 428, method = "polychoric")

corplot <- ggcorrplot(correlations$rho, p.mat=p.mat$p, insig = "pch", hc.order = FALSE, outline.col = "white", type = "lower", legend.title = "Correlation",sig.level = 0.05, pch.cex=1.2,
                      # insig = "blank",
           ggtheme = ggplot2::theme_minimal(),tl.srt = 90,colors = c("#E64B35FF", "white", "#3C5488FF") )+
  theme( panel.grid.major = element_blank(),
         axis.text.x = element_text(size = 6),
         axis.text.y = element_text(size = 6),
         panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))
#to show pvalue in the fig add p.mat=p.mat
# ggsave("Figures/corplot.png",corplot, width = 12, height = 8.5, dpi = 600, bg ="white")

```

```{r ParallelEFA, eval=FALSE, include=FALSE}
#This chunk holds code for factor identification method : Horn
# Parallel analysis
Horn <- paran(correlations$rho, iterations=500, centile=0, quietly=FALSE,
           status=TRUE, all=FALSE, cfa=FALSE, graph=TRUE,
           color=TRUE, col=c("black","red","blue"),
           lty=c(1,2,3), lwd=1, legend=TRUE, 
           width=640, height=640, grdevice=png, seed=0, mat=NA, n=NA)

Adjusted.EV <- data.frame(Horn$AdjEv)
Adjusted.EV$type = c('Adjusted Ev')
Adjusted.EV$num = c(row.names(Adjusted.EV))
Adjusted.EV$num = as.numeric(Adjusted.EV$num)
colnames(Adjusted.EV) = c('eigenvalue', 'type', 'num')

Random.EV <- data.frame(Horn$RndEv)
Random.EV$type = c('Random EV')
Random.EV$num = c(row.names(Random.EV))
Random.EV$num = as.numeric(Random.EV$num)
colnames(Random.EV) = c('eigenvalue', 'type', 'num')

Unadjusted.EV <- data.frame(Horn$Ev)
Unadjusted.EV$type = c('Unadjusted EV')
Unadjusted.EV$num = c(row.names(Unadjusted.EV))
Unadjusted.EV$num = as.numeric(Unadjusted.EV$num)
colnames(Unadjusted.EV) = c('eigenvalue', 'type', 'num')

parallel = ggplot(Adjusted.EV, aes(x=num, y=eigenvalue, shape=type, color=type)) +
  #Add lines connecting data points
  geom_line(colour ='#DC000099')+
    #geom_line(data = Random.EV) +
  #geom_line(data = Unadjusted.EV)+
    #Add the data points.
  geom_point(size=1.5,colour ='#3C5488FF')+
  #geom_point(data = Random.EV, size=3)+
  #geom_point(data = Unadjusted.EV, size=3)+
  scale_y_continuous(name="Eigen Value", limits=c(0, 7), breaks=seq(0,7,1)) +
   #Label the x-axis 'Factor Number', and ensure that it ranges from 1-max # of factors, increasing by one with each 'tick' mark.
  scale_x_continuous(name='Factor Number (Parallel Analysis)', limits=c(0, 48),breaks=seq(0,48,4))+
    #Add vertical line indicating parallel analysis suggested max # of factors to retain
  geom_hline(yintercept=1, linetype = 'dashed') + apatheme +theme(,legend.position = "None") 
#text = element_text(size = 25)

```

```{r ScreePlot, include=FALSE}

#This chunk holds code for factor identification method : Scree Plot

Scree <- scree(correlations$rho,factors=TRUE,pc=TRUE,main="(B)",
      hline=NULL,add=F)

FA.Scree <- data.frame(Scree$fv)
FA.Scree$type = c('Factor Analysis')
FA.Scree$num = c(row.names(FA.Scree))
FA.Scree$num = as.numeric(FA.Scree$num)
colnames(FA.Scree) = c('eigenvalue', 'type', 'num')

PA.Scree <- data.frame(Scree$pcv)
PA.Scree$type = c('Principal Component Analysis')
PA.Scree$num = c(row.names(PA.Scree))
PA.Scree$num = as.numeric(PA.Scree$num)
colnames(PA.Scree) = c('eigenvalue', 'type', 'num')

scree.plot <-ggplot(FA.Scree, aes(x=num, y=eigenvalue,color=type, shape=type))+
   geom_line(colour ='#DC000099')+
  #geom_line(data = PA.Scree)+ 
  geom_point(size=1.5,colour ='#3C5488FF')+
  #geom_point(data = PA.Scree, size=3)+
  scale_y_continuous(name="Eigen Value", limits=c(0, 5), breaks=seq(0,5,1)) +
  scale_x_continuous(name="Factor Number", limits=c(0, 18),breaks=c(0,6,18))+
  geom_hline(yintercept=1, linetype = 'dashed') + apatheme +theme(legend.position = "None")
#text = element_text(size = 25

```

```{r HullEFA, include=FALSE}
#This chunk holds code for factor identification method : Hull

# HULL
EFA.MRFA::hullEFA(EFA.data,extr = "ULS", index_hull = "CAF", display = TRUE, graph = T,
        details = TRUE)
hull <- ggplot2::last_plot()
Hull <- hull+ ggtitle(NULL)+xlab("Factor Number")+ylab("CAF")+ aes(color = '#DC000099')+geom_point(color ='#3C5488FF',size =1)+
  apatheme +theme(legend.position = "None") +scale_fill_identity()+
  scale_x_continuous(breaks=c(0,5,12))

```

```{r MAPefa, include=FALSE}
#This chunk holds code for factor identification method : MAP
#MAP
map <- psych::VSS(EFA.data, rotate = "varimax", fm = 'minres', n.obs =428 )
map.map <- as.data.frame(map$map)
colnames(map.map) <- "MAP Statistic"
map.statistics <- map$vss.stats[,c(1,2,5,6,7,10,11)]
full.map <- cbind(map.map,map.statistics)

write.csv(full.map, "Table_raw/map_stat.csv")
```

Inspection via the Scree plot ( Figure \@ref(fig:efa-plot-print)-B) suggested a six-factor solution, whereas the minimum average partial (MAP) method [@velicerDeterminingNumberComponents1976] (**Supplementary Table 3**) and Hull method [@lorenzo-sevaHullMethodSelecting2011] ( Figure \@ref(fig:efa-plot-print)-C) implied a five-factor solution for the LEBA questionnaire. As a result, we tested both five-factor and six-factor solutions.

```{r EFA, include=FALSE}
#This chunk holds code for five factor EFA (pa-varimax)

fa.5F.1 <- fa(r=correlations$rho, nfactors = 5, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)
AA <- print(fa.5F.1, cut = .3, digits = 3, sort = TRUE)

reduced.model.5F.1 <- dplyr::select(EFA.data ,
                                    -c( item20, item19, item05, item31, item44, item24, item43, item39, item22, item02, item18, item48, item42, item29, item06, item15, item23, item28, item14))

correlations.red.5F.1 <- polychoric(reduced.model.5F.1,correct = 0)

fa.5F.2 <- fa(r=correlations.red.5F.1$rho, nfactors = 5, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428)

BB <- print(fa.5F.2, cut = .3, digits = 3, sort = TRUE)

reduced.model.5F.2 <- dplyr::select(EFA.data, -c( item20, item19, item05, item31, item44, item24, item43, item39, item22, item02, item18, item48, item42, item29, item06, item15, item23, item28, item14, item34, item47, item21, item13))

correlations.red.5F.2 <- polychoric(reduced.model.5F.2,correct = 0)

fa.5F.3 <- fa(r=correlations.red.5F.2$rho, nfactors = 5, fm= "pa",rotate ="varimax",
              residuals = TRUE, SMC = TRUE, n.obs =428,max.iter = 500 )

CC <- print(fa.5F.3, cut = .3, digits = 3, sort = TRUE)

var1 <- CC$Vaccounted[2,1]*100
var2 <- CC$Vaccounted[2,2]*100
var3 <- CC$Vaccounted[2,3]*100
var4 <- CC$Vaccounted[2,4]*100
var5 <- CC$Vaccounted[2,5]*100

omega <- omega(reduced.model.5F.2,nfactors=5,fm="minres",n.iter=1,p=.05,poly=T,key=NULL,
    flip=TRUE,digits=2, title="Omega",sl=TRUE,labels=NULL,
    plot=TRUE,n.obs=NA,rotate="oblimin",Phi=NULL,option="equal",covar=FALSE)
```

```{r Cronbach alpha for 5 fators, message=TRUE, warning=FALSE, include=FALSE}
#This chunk holds code for estimating cronbach's alpha for the five factors.

Five.F1 <- dplyr::select(EFA.data, item16, item36, item17)
Five.F2 <- dplyr::select(EFA.data, item11, item10, item12, item07, item08, item09) 
Five.F3 <-  dplyr::select(EFA.data, item27, item03,item40, item30, item41)
Five.F4 <-  dplyr::select(EFA.data, item33,item32,item35, item37, item38)
Five.F5 <- dplyr::select(EFA.data, item46, item45, item25, item04, item01, item26)

#ordinal alpha
Five.F1.alpha <- psych::alpha(polychoric(Five.F1)$rho,check.keys=TRUE)
Five.F2.alpha <- psych::alpha(polychoric(Five.F2)$rho,check.keys=TRUE)
Five.F3.alpha <- psych::alpha(polychoric(Five.F3)$rho,check.keys=TRUE)
Five.F4.alpha <- psych::alpha(polychoric(Five.F4)$rho,check.keys=TRUE)
Five.F5.alpha <- psych::alpha(polychoric(Five.F5)$rho,check.keys=TRUE)

F1.alpha <- Five.F1.alpha$total$raw_alpha
F2.alpha <- Five.F2.alpha$total$raw_alpha
F3.alpha <- Five.F3.alpha$total$raw_alpha
F4.alpha <- Five.F4.alpha$total$raw_alpha
F5.alpha <- Five.F5.alpha$total$raw_alpha

alpha.tab <- rbind(F1.alpha,F2.alpha,F3.alpha,F4.alpha,F5.alpha )

```

```{r tabledown, include=F}
#This chunk holds codes for creating factor analysis output table

efa.table <- tabledown::fac.tab(fa.5F.3,cut=.3, complexity = F)
efa.table$Communality <- round(as.numeric(efa.table$Communality),2)

item16 <- Hmisc::label(EFA.data$item16)
item36 <- Hmisc::label(EFA.data$item36)
item17 <- Hmisc::label(EFA.data$item17)
item11 <- Hmisc::label(EFA.data$item11)
item10 <- Hmisc::label(EFA.data$item10)
item12 <- Hmisc::label(EFA.data$item12)
item07 <- Hmisc::label(EFA.data$item07)
item08 <- Hmisc::label(EFA.data$item08)
item09 <- Hmisc::label(EFA.data$item09)
item27 <- Hmisc::label(EFA.data$item27)
item03 <- Hmisc::label(EFA.data$item03)
item40 <- Hmisc::label(EFA.data$item40)
item30 <- Hmisc::label(EFA.data$item30)
item41 <- Hmisc::label(EFA.data$item41)
item33 <- Hmisc::label(EFA.data$item33)
item32 <- Hmisc::label(EFA.data$item32)
item35 <- Hmisc::label(EFA.data$item35)
item37 <- Hmisc::label(EFA.data$item37)
item38 <- Hmisc::label(EFA.data$item38)
item46 <- Hmisc::label(EFA.data$item46)
item45 <- Hmisc::label(EFA.data$item45)
item25 <- Hmisc::label(EFA.data$item25)
item04 <- Hmisc::label(EFA.data$item04)
item01 <- Hmisc::label(EFA.data$item01)
item26 <- Hmisc::label(EFA.data$item26)

efa.stems <- as.data.frame(rbind(item16,item36, item17, item11, item10, item12, item07, item08, item09, item27, item03, item40, item30, item41, item33, item32, item35, item37, item38, item46, item45, item25, item04, item01, item26, ""))
colnames(efa.stems) <- "Stem"

efa.tab.ful <- cbind(efa.table[,1],efa.stems, efa.table[,2:7] )

efa.tab.ful[26,8] <- "-"

#Communality
Fac.5 <- fa.5F.3
min.com <- min(Fac.5$communality)
max.com <- max(Fac.5$communality)

# loading
Fac.5.min.max <- Fac.5$loadings %>% 
  subset(abs(Fac.5$loadings)>(.3))
min.loadings <- min(abs(Fac.5.min.max))
max.loadings <- max(Fac.5.min.max)

# Residuals
residual.5F =residuals(fa.5F.3, diag = FALSE, na.rm =T)
#Count of numbers of residuals>.05
BigRR= sum(residual.5F>abs(.05), na.rm =T)
print(BigRR)

#Total number of off diagonal elements in the data matrix
totRR = length(reduced.model.5F.2)*(length(reduced.model.5F.2)-1)/2

# Proportion of off-diagonal elements >.05
sumR <- sum(BigRR/totRR*100)

```

```{r EFARes, message=FALSE, warning=FALSE, include=F}
#This chunk holds codes for creating the residual plot for 5 factor EFA (pa-varimax)
res.dat <- as.data.frame(unclass(residual.5F))
residual.data <- as.data.frame(gather(res.dat, Items, Residuals))

residual.plot <- ggplot(residual.data, aes(x=Residuals)) + geom_histogram(binwidth=.04, col=I("black"),fill="#00A08799") + xlim(-.2,.2)+ ylim(0,250)+apatheme +geom_vline(xintercept = .05)+theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12))+
  labs( y="Frequency")

```

```{r efa-plot, include=F}
#This chunk holds code for compiled factor identification plots

cor.scree <- cowplot::plot_grid(corplot, scree.plot,
                                labels = c("A", "B"),
                                nrow =1, ncol =2,
                                rel_widths = c(2, 1),
                                label_fontfamily = "Arial",
                  label_fontface = "plain")

hull.residual <- cowplot::plot_grid(Hull, residual.plot,
                                labels = c("C", "D"),
                                nrow =1, ncol =2,
                                                                label_fontfamily = "Arial",
                  label_fontface = "plain")

efa.plot <- cowplot::plot_grid(cor.scree, hull.residual,

                                nrow =2)

# efa.plot <- cowplot::plot_grid(corplot , scree.plot,Hull, residual.plot,
#                    labels = "AUTO",
#                    ncol=2,nrow=2, 
#                   align = "v", 
#                   label_fontfamily = "Arial",
#                   label_fontface = "plain")
ggsave("Figures/Figure4.png",efa.plot, width = 7, height = 8, dpi = 600)
```

(ref:efa-plot-print) (A) Inter-item polychoric correlation coefficients for the 48 items. 4.9 % inter-item correlation coefficients were higher than \|.30\|. 'x' denotes non-significant item-total correlation. (B) The Scree plot suggested six factors. (C) Hull method indicated that five factors were required to balance the model fit and number of parameters. (D) The histogram of nonredundant residual correlations indicated that 26% of inter-item correlations were higher than .05, hinting at a possible under-factoring.

```{r efa-plot-print, fig.align= 'center', fig.cap='(ref:efa-plot-print)',  warning=FALSE,  out.height='60%', out.width='110%'}

knitr::include_graphics('Figures/Figure4.png', dpi =600)

```

Applying varimax rotation, we conducted three rounds of EFA with the initial 48 items and gradually discarded problematic items (cross-loading items and items with factor loading \<.30).
Finally, a five-factor EFA solution with 25 items was accepted with all factor-loading higher than .30 and no cross-loading greater than .30. Table \@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute values of the factor-loadings ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` and `r apa(max.com, 2, F)`. However, the histogram of the absolute values of nonredundant residual correlations (Figure \@ref(fig:efa-plot-print)-D) displayed that `r apa(sumR,0, T)`% of correlations were greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018]. Subsequently, we fitted a six-factor solution, wherefrom a factor with only two salient variables emerged, thus disqualifying the six-factor solution (**Supplementary Table 4**).

In the five-factor solution, the first factor contained three items and explained `r apa(var1,2,F)`% of the total variance with an internal reliability coefficient ordinal $\alpha$ = `r apa(F1.alpha,2,F)`. All the items in this factor encapsulated the individual's preference for using blue light filters in different light environments. The second factor contained six items and explained `r apa(var2,2,F)`% of the total variance with an internal reliability coefficient ordinal $\alpha$ = `r apa(F2.alpha,2,F)`. Items under this factor incorporated the individuals' hours spent outdoor. The third factor contained five items and explained `r apa(var3,2,F)`% of the total variance. Items under this factor covered the specific behaviours of using a phone and smart-watch in bed. The internal consistency reliability coefficient was ordinal $\alpha$ = `r apa(F3.alpha,2,F)`. The fourth factor comprised five items and explained `r apa(var4,2,F)`% of the total variance with an internal consistency coefficient, ordinal $\alpha$ = `r apa(F4.alpha,2,F)`. These five items investigated the behaviours related to the individual's light exposure before bedtime. The fifth factor encompassed six items and explained `r apa(var5,2,F)`% of the total variance. This factor captured the individual's morning and daytime light exposure-related behaviour. The internal consistency reliability yielded ordinal $\alpha$ = `r apa(F5.alpha,2,F)`.

Finally, we examined the factor's interpretability in the five-factor solution and weighed it against the psychometric properties as we considered it essential to attain a balance between the two. As we deemed the five derived factors interpretable and relevant concerning our aim to capture light exposure-related behaviour, we retained all of them with 25 items for our confirmatory factor analysis (CFA), despite the apparent low reliability of the fifth factor. Two of the items showed negative factor-loading (items 44 and 21). Upon re-inspection, we recognized these items to be negatively correlated to the respective factor, and thus, we reverse-scored these two items in the CFA analysis. The internal consistency coefficient McDonald's $\omega_t$ for the total scale was `r apa(omega$omega.tot,2,T)`.

```{r EFATable, results='asis'}

apa_table(efa.tab.ful, caption = "Factor loadings and communality of the retained items in EFA using principal axis extraction method (n=482).", note = "Only loading > .30 is reported.", 
          font_size = "footnotesize", escape=F,landscape = T,
           align = c("p{1cm}","p{10cm}", rep("p{1cm}", ncol(efa.tab.ful))))
```

```{r EFAplotprep, eval=FALSE, include=FALSE}
#This chunk holds codes for co-plotting EFA-plot and residual plot.

library(jpeg)
library(png)
library(cowplot)

img1 <- png::readPNG("Figures/EFAplot.png")

p1 <- ggplot2::ggplot() + ggplot2::annotation_custom(grid::rasterGrob(img1,
                                                                      width=.4,
                                                                      height=.8),
                                                     -Inf, Inf, -Inf, Inf)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())

EFAplot.1 <- cowplot::plot_grid(
  NULL, residual.plot,
  labels = c("", "B"),
  align="h",
  ncol = 1,
  label_size = 15,
  label_fontfamily = "Arial",
  label_fontface = "plain",
  label_y = .99
)

EFAplot.2 <- cowplot::plot_grid(
  p1, EFAplot.1,
  labels = c("A",""),
  align="h",
  ncol = 2,
  label_size = 15,
  label_fontfamily = "Arial",
  label_fontface = "plain",
  label_y = .99,
  label_x = .02
)

ggsave("Figures/EFAfull.png",EFAplot.2, width = 6, height =7, dpi = 1200)

```

### Confirmatory Factor Analysis

```{r gt-CFA, include =F}
gt.cfa.data <- gt.data[429:690,]
gt.cfa.data.table <- gt.cfa.data %>% 
  dplyr::select(item16, item36, item17,
                 item11, item10,item12,item07,item08,item09,
                 item27,  item03, item40,  item30, item41,
                 item32,item35,item38, item33, 
                 item46, item45, item25, item04, item01 )

gt.cfa.long <- gather(gt.cfa.data.table, Items, value)
gt.cfa.long$value <- as.numeric(as.character(gt.cfa.long$value))
gt.cfa.tab <- gt.cfa.long %>% 
  group_by(Items) %>% 
# calculate summary stats & create data for the histogram and density plot
  dplyr::summarise(
    nr = n(),
    mean = mean(value, na.rm = TRUE),
    med = median(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    hist_data = list(value),
    dens_data = list(value),
    .groups = "drop"
  ) %>%

  gt()

#Preparation for likert data
cfa.LEBA.likert <- as.data.frame(gt.cfa.data.table)

recod_LEBA <- c( "1" = "Never", "2" = "Rarely", "3"= "Sometimes","4" = "Often",
                 "5" = "Always")
cfa.LEBA.likert <-  mutate(cfa.LEBA.likert, across(starts_with("item"), ~unname(recod_LEBA[.])))

cfa.LEBA.Factor = as.data.frame(lapply(cfa.LEBA.likert,factor,
                                   ordered = T))
#get the items name
cfa.items <- names(cfa.LEBA.Factor) 
#Calculate percentage
cfa.percentage <- kutils::likert(cfa.LEBA.Factor, vlist = cfa.items )

cfa.percentage <- cfa.percentage$table %>% 
  as.data.frame(.)

#data wrangling
labels <- c("Always", "Never","Often", "Rarely", "Sometimes","Total")
as.data.frame(labels)

full.cfa.percentage <- cbind(labels,cfa.percentage) #tables with labels  
full.cfa.percentage <- t(full.cfa.percentage ) #transpose
as.data.frame(full.cfa.percentage)
full.cfa.percentage1 <- full.cfa.percentage[-1,-6] #removing 1st row and total column
full.cfa.percentage2 <- full.cfa.percentage1[, c(2, 4, 5, 3,1)]# rearranging
as.data.frame(full.cfa.percentage2)

colnames(full.cfa.percentage2) <- c("Never","Rarely","Sometimes","Often","Always")

cfa.Items <- rownames(full.cfa.percentage2)
as.data.frame(cfa.Items)
full.cfa.percentage3 <- cbind(cfa.Items,full.cfa.percentage2)

full.cfa.percentage3 <- full.cfa.percentage3[order(cfa.Items),] 
full.cfa.percentage3 <- as.data.frame(full.cfa.percentage3[,-1]) %>% 
  gt()

gt.cfa.table <- gt.cfa.tab$'_data'
gt.liket.cfa <- full.cfa.percentage3$"_data"

cfaTable <-  cbind( gt.cfa.table, gt.liket.cfa) %>% 
  gt()

cfa.gt <- cfaTable$"_data"

```

```{r cfagt,include=F}
#This chunk holds codes for creating 'gtExtra' based descriptive figures (CFA Descriptives Supplementary

cfa.gt %>% 
gt() %>% 
# histogram and density plots
  gtExtras::gt_sparkline(
    hist_data,
    type = "histogram",
    line_color = "black", 
    fill_color = "#00A08799",
    bw = 1,
    same_limit = TRUE)%>%
  gtExtras::gt_sparkline(
    dens_data,
    type = "density",
    line_color = "black", 
    fill_color = "#E64B3599",
    bw = 0.75,
    same_limit = TRUE
  )%>%
# format decimals
  fmt_number(columns = mean:sd, decimals = 1) %>%
# header
  tab_header(
    title = md("**Light Exposure Behavior Assessment**"),
    subtitle = md("Summary Descriptives CFA Sample (n=262)")
  ) %>% 
  # tab_footnote(
  #   footnote = md("**Histogram**"),
  #   locations = cells_column_labels(columns = hist_data)
  # ) %>%
  # tab_footnote(
  #   footnote = md("**Density**"),
  #   locations = cells_column_labels(columns = dens_data)
  # ) %>% 
#create groups of columns
tab_spanner(
  label = "Items",
  columns = Items
) %>%
  tab_spanner(
    label = "Summary Statistics",
    columns = nr:sd
  ) %>%
  tab_spanner(
    label = "Graphics",
    columns = hist_data:dens_data
  ) %>% 
  tab_spanner(
    label = "Response Pattern",
    columns = Never:Always
  ) %>%
# change column names to appear in the table
cols_label(
  Items = (""),
  nr = ("n"),
  mean = ("Mean"),
  med = ("Median "),
  sd = (("SD")),
  hist_data = "Histogram",
  dens_data = "Density"
) %>%
# set alignment as per wish
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_align_table_header(align = "left") %>%
# add coloured dots and lines on the first column
  gt_plt_dot(
    mean,
    Items,
    palette = "ggthemes::fivethirtyeight") %>% 
  tab_row_group(
      label = md("**F5: Using light in the morning and during daytime**"),
      rows = c(1,3,12,13,22,23) 
        ) %>% 
  tab_row_group(
      label = md("**F4: Using light before bedtime**"),
      rows = c(15,16,17,19) 
        ) %>% 
  tab_row_group(
      label = md("**F3: Using phone and smart-watch in bed**"),
      rows = c(2,13,14,20,21) 
        ) %>% 
  tab_row_group(
      label = md("**F2: Spending time outdoors**"),
      rows = c(4,5,6,7,8,9) 
        ) %>% 
  tab_row_group(
      label = md("**F1: Wearing blue light filters**"),
      rows = c(10,11,18) 
        ) %>% 
  gtsave("Figures/S2_Fig.png",vwidth = 7000)

```

```{r CFAdataframe, include=FALSE}
#This chunk holds codes for CFA data preparation.

### CFA Data
CFA.data <- sem.data[429:690,]

## Recoding CFA.data$item8 into CFA.data$Ritem8
CFA.data$Ritem08 <- as.character(CFA.data$item08)
CFA.data$Ritem08 <- fct_recode(CFA.data$Ritem08,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)
CFA.data$Ritem08 <- as.numeric(as.character(CFA.data$Ritem08))

## Recoding CFA.data$item37 into CFA.data$item37_rec
CFA.data$Ritem37 <- as.character(CFA.data$item37)
CFA.data$Ritem37 <- fct_recode(CFA.data$Ritem37,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)
CFA.data$Ritem37 <- as.numeric(as.character(CFA.data$Ritem37))

## Recoding CFA.data$item26 into CFA.data$Ritem26
CFA.data$Ritem26 <- as.character(CFA.data$item26)
CFA.data$Ritem26 <- fct_recode(CFA.data$Ritem26,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)

#Hitogram
#psych::multi.hist(CFA.data[,sapply(CFA.data, is.numeric)])
```

```{r CFA25items, include=FALSE }
#This chunk holds codes for 1st CFA

LB.model.Cor.1 <- "F1 =~ a1*item16 + item36 + a2*item17
                 F2 =~ item11 + item10 + item12+ item07+ Ritem08+ item09
                 F3 =~  item27+ item03+ item40 + item30 + item41
                 F4 =~ a3*item32 + item35 + item38+ a4*item33 + item37
                 F5 =~ item46+ item45+ item25+ item04+ item01 +item26
                       a1==a2
                       a3==a4"

fit.Cor.1 <- cfa(LB.model.Cor.1, data = CFA.data, ordered = names(CFA.data),estimator = "WLSMV")

## Summary of Model 
cfa.sum.1 <- summary(fit.Cor.1, fit.measures =TRUE,standardized = TRUE,rsq =TRUE)

## Selected Fit measures 
fit.1 <- fitmeasures (fit.Cor.1,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

reliability1 <- semTools::reliability(fit.Cor.1)

```

```{r mod1, eval=FALSE, include=FALSE}
#This chunk holds codes for modification indices for 1st CFA
modfit.Cor.one <- modindices(fit.Cor.1, sort. = TRUE) 
modfit.Cor.one[modfit.Cor.one$mi>3.84,]
```

```{r CFA23withMI, include=FALSE}
#This chunk holds codes for 2nd CFA (Accepted)

LB.model.Cor.2 <- "F1 =~ a1*item16 + item36 + a2*item17
                 F2 =~ item11 + item10 + item12+ item07+ Ritem08+ item09
                 F3 =~  item27+ item03+ item40 + item30 + item41
                 F4 =~ a3*item32 + item35 + item38+ a4*item33 
                 F5 =~ item46+ item45+ item25+ item04+ item01 
                       a1==a2
                       a3==a4
                item30 ~~  item41
                     F1 ~~ 0*F2
F1 ~~ 0*F3
F1 ~~ 0*F4
F1 ~~ 0*F5
F2 ~~ 0*F3
F2 ~~ 0*F4
F2 ~~ 0*F5
F3 ~~ 0*F4
F3 ~~ 0*F5
F4 ~~ 0*F5"

#item26 & 37 are removed

## Fit the model 
fit.Cor.2 <- cfa(LB.model.Cor.2, data = CFA.data, ordered = names(CFA.data),estimator = "WLSMV")

rel.k <- semTools::reliability(fit.Cor.2, return.total = T)
## Summary of Model 
cfa.sum.2 <- summary(fit.Cor.2, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

## Selected Fit measures 
fit.2 <- fitmeasures (fit.Cor.2,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

```

```{r mod2, eval=FALSE, include=FALSE}
#This chunk holds codes for modification indices for 2nd CFA
modfit.Cor.one <- modindices(fit.Cor.2, sort. = TRUE) 
modfit.Cor.one[modfit.Cor.one$mi>3.84,]
```

Table \@ref(tab:tabCfa) compares the CFA fit indices of the original CFA five-factor model with 25 and the post-hoc modified model with 23 items, respectively. The 25-item model attained an acceptable fit (CFI =`r apa(fit.1[5],2,F)`; TLI = `r apa(fit.1[6],2,F)`; RMSEA = `r apa(fit.1[7],2,F)` [`r apa(fit.1[8],2,F)`-`r apa(fit.1[9],2,F)`, 90% CI]) with two imposed equity constraints on item pairs 32-33 [item 32: I dim my mobile phone screen within 1 hour before attempting to fall asleep; item 33: I dim my computer screen within 1 hour before attempting to fall asleep] and 16-17 [item 16: I wear blue-filtering, orange-tinted, and/or red-tinted glasses indoors during the day; item 17: I wear blue-filtering, orange-tinted, and/or red-tinted glasses outdoors during the day]. Item pair 32-33 describes the preference for dimming the electric devices' brightness before bedtime, whereas item pair 16-17 represents the preference for using blue filtering or coloured glasses during the daytime. Given the similar nature of captured behaviours within each item pair, we accepted the imposed equity constraints. Nevertheless, the SRMR value exceeded the guideline recommendation (SRMR = `r apa(fit.1[10],2,F)`).

In order to improve the model fit, we conducted a post-hoc model modification. Firstly, the  modification indices suggested cross-loadings between item 37 and 26 [item 37: I purposely leave a light on in my sleep environment while sleeping; item 26: I turn on my ceiling room light when it is light outside], which were hence discarded.  Secondly, items 30 and 41 [item 30: I look at my smart-watch within 1 hour before attempting to fall asleep; item 41: I look at my smart-watch when I wake up at night] showed a tendency to co-vary in their error variance (MI = 141.127, p<.001 ). By allowing the latter pair of items (30 $\&$ 41) to co-vary, the model's error variance attained an improved fit (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)` [`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]; SRMR = `r apa(fit.2[10],2,F) `).  Internal consistency ordinal $\alpha$ for the five factors of the LEBA were `r apa(rel.k[2,1],2,F)`, `r apa(rel.k[2,2],2,F)`, `r apa(rel.k[2,3],2,F)`, `r apa(rel.k[2,4],2,F)`, `r apa(rel.k[2,5],2,F)`, respectively.

Accordingly, we accept the 5-factor model with 23 items, finalizing the long Form of LEBA (see **Supplementary File 1**). The Internal consistency McDonald's $\omega_t$ coefficient for the total scale yielded `r apa(rel.k[5,6],2,F)`. Figure \@ref(fig:figcfa) depicts the obtained CFA structure, while  **Supplementary Figure 2** depicts the data distribution and endorsement pattern of the retained 23 items in our CFA sample.

```{r cfaMat, results='asis'}
#This chunk holds codes to create CFA output table (better options are avaiable at 'tabledown' package)

CFa.matrix <- matrix("NA", nrow =2, ncol = 9)
colnames(CFa.matrix) = c("Model", "$\\chi^{2}$", "df", "CFI", "TLI", "RMSEA", "RMSEA 90\\% Lower CI", "RMSEA 90\\% Upper CI", "SRMR")
CFa.matrix[1,] = c("Model 1", 
                   apa(cfa.sum.1$FIT[6],2,T), 
                   apa(cfa.sum.1$FIT[7],2,T), 
                   apa(cfa.sum.1$FIT[19],2,F), 
                   apa(cfa.sum.1$FIT[20],2,T),
                   apa(cfa.sum.1$FIT[27],2,T),
                   apa(cfa.sum.1$FIT[28],2,T),
                   apa(cfa.sum.1$FIT[29],2,T),
                   apa(cfa.sum.1$FIT[35],2,T))

CFa.matrix[2,] = c("Model 2", 
                   apa(cfa.sum.2$FIT[6],2,T), 
                   apa(cfa.sum.2$FIT[7],2,T), 
                   apa(cfa.sum.2$FIT[19],2,F), 
                   apa(cfa.sum.2$FIT[20],2,T),
                   apa(cfa.sum.2$FIT[27],2,T),
                   apa(cfa.sum.2$FIT[28],2,T),
                   apa(cfa.sum.2$FIT[29],2,T),
                   apa(cfa.sum.2$FIT[35],2,T))
```

```{r tabCfa, results='asis'}
apa_table(CFa.matrix, caption = "Confirmatory Factor Analysis model fit indices of the two model: (a) Model 1: five factor model with 25 items (b) Model 2:  five factor model with 23 items. Model 2 attained the best fit.", align = "c", landscape = T, escape=F, note = "df: Degrees of Freedom; CFI: Comparative Fit Index; TLI: Tucker Lewis Index; RMSEA: Root Mean Square Error of Approximation; CI: Confidence Interval; SRMR: Standardized Root Mean Square.")
```

```{r cfaplot, eval=FALSE, include=FALSE}
#This chunk holds codes to store CFA plot

png(filename="Figures/CFAplot2.png", 
    type="cairo",
    units="in", 
    width=12, 
    height=12, 
    pointsize=12, 
    res=1200)
semPaths (fit.Cor.2 , 
          what= "std",
          #"hide", #(hides coefficeits)
          #whatLabels = "std",
          intercepts = F,
          style ="OpenMx",
          #residScale = 6,
          theme = "colorblind",
          nCharNodes = 0,
          reorder =T,
          rotation =2,
          layout ="tree",
          cardinal = T,
          curvePivot =T,
          sizeMan =8,#items length
          sizeMan2 = 2,#items
          sizeLat = 12,#factors
          thresholds = FALSE,
          equalizeManifests =F,
          fade = FALSE,
          edge.label.cex = .8,
          #exoCov = T,
          centerLevels = T,
          #edge.color="black",
          label.scale=T,
          label.cex=1.2, #Font of factor and item name
          residuals=T,
           #exoVar=FALSE,
          #fixedStyle=6, #Style of arrow (guide item)
          #freeStyle=1#Style of arrow (other item), #XKCD = TRUE

          )

dev.off()
```

```{r figcfa, fig.cap = "Five factor  model of LEBA obtained by confirmatory factor analysis. By allowing item pair 41 and 30 to co-vary their error variance our model attained the best fit.",  out.height='100%',out.width='100%', warning=FALSE}

knitr::include_graphics("Figures/Figure5.png", dpi =600)

```

### Measurement Invariance

```{r InvarianceAnalysis_data, include=FALSE}
#This chunk holds codes to prepare measurement invariance data.

invariance.data <- cbind(CFA.data, data$slypos_demographics_language.factor[429:690])
colnames(invariance.data)[52] <- "English_Speaking"

mi.count <- invariance.data %>% 
  group_by(English_Speaking) %>% 
  count() %>% 
  as.data.frame()

mi.descriptives <- CFA.descriptives[, c(1,2,5,6,7)] 
colnames(mi.descriptives)[3] <- "English_Speaking"

english.speaking <- mi.descriptives %>% 
  subset(English_Speaking=="Yes")

non.english.speaking <- mi.descriptives %>% 
  subset(English_Speaking=="No")

mi.des.sum <- mi.descriptives %>% 
psych::describeBy("English_Speaking")

mi.english.sum <- english.speaking %>% 
psych::describeBy("slypos_demographics_sex.factor")

mi.non.english.sum <- non.english.speaking %>% 
psych::describeBy("slypos_demographics_sex.factor")

mi.des.sum.english <- as.data.frame(mi.des.sum$Yes)
mi.des.sum.nonenglish <- as.data.frame(mi.des.sum$No)
```

Our CFA sample consisted of `r mi.count[1,2]` native English speakers and `r mi.count[2,2]` non-native English speakers, whose demographic data are contrasted in **Supplementary Table 5**. As shown in Table \@ref(tab:InvarianceTab), the employed five-factor model generated acceptable fit indices over all of the fitted MI models. The model fit did not significantly decrease across the nested models implying the acceptability of the highest measurement invariance model (residual model).

```{r Invariancedetail, include=FALSE}
#This chunk holds codes for measurement Invariance

LB.model.Cor.2 <- "F1 =~ a1*item16 + item36 + a2*item17
                 F2 =~ item11 + item10 + item12+ item07+ Ritem08+ item09
                 F3 =~  item27+ item03+ item40 + item30 + item41
                 F4 =~ a3*item32 + item35 + item38+ a4*item33 
                 F5 =~ item46+ item45+ item25+ item04+ item01 
                       a1==a2
                       a3==a4
                item30 ~~  item41
                       "  
# Configural invariance ####
configural <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV")

summary(configural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (configural,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

#Metric
weak <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV", 
            group.equal = "loadings")

summary(weak, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

fitmeasures (weak,c("gfi", "agfi", "nfi","rfi", 
                       "cfi.scaled","tli.scaled",
                       "rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled","srmr"))

First.comp <- compareFit (configural, weak)
summary(First.comp)

#Scaler
strong <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV", 
              group.equal = c("loadings", "intercepts"))
summary(strong, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

Second.comp <- compareFit  (weak, strong)
summary(Second.comp)

#residual
strict <- cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV",  
              group.equal = c("loadings", "intercepts", "residuals"))

summary(strict, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
residual.fitm <- as.data.frame(fitmeasures (strict,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic")))

Third.comp <- compareFit(strict,strong)
summary(Third.comp)
#structural
structural <-cfa(model = LB.model.Cor.2,
                  data = invariance.data,
                  group = "English_Speaking",
                  ordered =   names(CFA.data),
                  estimator = "WLSMV",  
                 group.equal = c("loadings", "intercepts", "residuals", "means","lv.variances","lv.covariances"))

summary(structural, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)
fitmeasures (structural,c("gfi", "agfi", "nfi","rfi", "cfi","tli","rmsea", "srmr","aic"))

fourth.comp <- compareFit  (structural,strict)
summary(fourth.comp)
comfit.par <- compareFit(configural, weak, strong, strict)
summary (comfit.par)

models <-  list("Configural" = configural, 
                "Metric" = weak, 
                "Scalar" = strong, 
                "Residual" = strict)

Invariance.table <- compareLavaan(models,
              nesting = "Configural > Metric > Scalar > Residual", 
              fitmeas = c("chisq.scaled", "df",  "cfi.scaled","tli.scaled","rmsea.scaled", "rmsea.ci.lower.scaled", "rmsea.ci.upper.scaled" ),
              scaled = T,
              chidif = T, digits = 2)

colnames(Invariance.table) <- c("$\\chi^{2}$", "df", "CFI", "TLI", "RMSEA", "RMSEA 90\\% Lower CI", "RMSEA 90\\% Upper", "$\\Delta$ $\\chi^{2}$", "$\\Delta$ df*", "p")

```

```{r InvarianceTab, results='asis'}
apa_table(Invariance.table, align = "c",  caption = "Measurement Invariance analysis on CFA sample (n=262) across native and non-native English speakers.", note = "df: Degrees of Freedom; CFI: Comparative Fit Index; TLI: Tucker Lewis Index; RMSEA: Root Mean Square Error of Approximation; CI: Confidence Interval; SRMR: Standardized Root Mean Square; a = Metric vs Configural; b = Scalar vs Metric; c = Residual vs Scalar; d = Structural vs Residual; * =  df of model comparison.", landscape = T, font_size = "footnotesize", escape=F )

```

## Secondary Analysis: Grade Level Identification and Semantic Scale Network Analysis
A grade level identification and Semantic Scale analysis were additionally administered to assess the LEBA's (23 items) language-based accessibility and its' semantic relation to other questionnaires. The results of the Flesch-Kincaid grade level analysis [@flesch1948new] displayed a required educational grade level of 3.33 with age above 8.33 years, implying that the LEBA instrument should be understandable for students of grade four at least 8.33 years old. Furthermore, the
Semantic Scale Network (SSN) analysis [@rosenbusch2020semantic] indicated that the LEBA appeared most strongly related to scales about sleep: The "Sleep Disturbance Scale For Children" [@bruni1996sleep] and the "Composite International Diagnostic Interview (CIDI): Insomnia"[@robins1988composite]. The cosine similarity yielded values between .47 to .51.

```{r textanalysis, include=F}
#This chunk holds codes for semantic analysis and readability test

#devtools::install_github("unDocUMeantIt/koRpus")
library(koRpus)# stable release
#koRpus::install.koRpus.lang(lang=c("en"))
library(koRpus)
library(koRpus.lang.en)
ques <- tokenize("Table_raw/items.txt", 
                 format = "file",
                 fileEncoding = NULL,
                 split = "[[:space:]]",
                 ign.comp = "-",
                 heuristics = "abbr",
                 heur.fix = list(pre = c("\u2019", "'"), suf = c("\u2019", "'")),
                 abbrev = NULL,
                 tag = TRUE,
                 lang = "en",
                 sentc.end = c(".", "!", "?", ";", ":"),
                 detect = c(parag = FALSE, hline = FALSE),
                 clean.raw = NULL,
                 perl = FALSE,
                 stopwords = NULL,
                 stemmer = NULL,
                 doc_id = NA,
                 add.desc = "kRp.env")

readability <- readability(ques,
                           hyphen = NULL,
                           index = "Flesch.Kincaid",
                           parameters = list(),
                           word.lists = list(Bormuth = NULL, Dale.Chall = NULL,                              Harris.Jacobson = NULL, Spache =NULL),
                           fileEncoding = "UTF-8",
                           sentc.tag = "sentc",
                           nonword.class = "nonpunct",
                           nonword.tag = c(),
                           quiet = FALSE,
                           keep.input = NULL,
                           as.feature = FALSE)
```

## Developing Short form of LEBA: IRT-Based Analysis
In order to derive a short form of the LEBA instrument, we fitted each factor of the LEBA with the graded response model [@samejima1997handbook] to the combined EFA and CFA sample (n=690). The resulting item discrimination parameters of the scale fell into categories of "very high" (10 items), "high" (4 items), "moderate" (4 items), and "low" ( 5 items), indicating a good range of discrimination along the latent trait level ($\theta$) (**Supplementary Table 6**). An examination of the item information curve (**Supplementary Figure 3**) revealed five items (1, 25, 30, 38, & 41) with relatively flat curves (I($\theta$) \<.20). We discarded those items, culminating in a short form of LEBA with five factors and 18 items (**Supplementary File 2**).

```{r IRTdata, include=FALSE}
#This chunk holds codes for preparing IRT data and 1st IRT model fitting

## Recoding sem.data$item8 into sem.data$Ritem8
sem.data$Ritem08 <- as.character(sem.data$item08)
sem.data$Ritem08 <- fct_recode(sem.data$Ritem08,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)
sem.data$Ritem08 <- as.numeric(as.character(sem.data$Ritem08))
# Recoding sem.data$item26 
sem.data$Ritem26 <- as.character(sem.data$item26)
sem.data$Ritem26 <- fct_recode(sem.data$Ritem26,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)
sem.data$Ritem26 <- as.numeric(as.character(sem.data$Ritem26))

# Recoding sem.data$item37 i
sem.data$Ritem37 <- as.character(sem.data$item37)
sem.data$Ritem37 <- fct_recode(sem.data$Ritem37,
  "5" = "1",
  "4" = "2",
  "2" = "4",
  "1" = "5"
)
sem.data$Ritem37 <- as.numeric(as.character(sem.data$Ritem37))

F1 <- sem.data %>% 
  dplyr::select(item16, item36,  item17)

F2 <- sem.data %>% 
  dplyr::select(item11,  item10 , item12, item07, Ritem08, item09)

F3 <- sem.data %>% 
 dplyr::select(item27, item03,  item40,  item30,  item41)

F4 <- sem.data %>% 
 dplyr::select(item32, item35,  item38,  item33)

F5 <- sem.data %>% 
  dplyr::select(item46, item45, item25, item04, item01 )

## F1 IRT model####
F1_fit <- mirt(F1, model = 1, itemtype = 'graded', 
               SE = TRUE, Se.type = 'MHRM',
               technical = list(NCYCLES = 10000))

## F2 IRT model####
F2_fit <- mirt(F2, model = 1, itemtype = 'graded', SE = TRUE,
               Se.type = 'MHRM',technical = list(NCYCLES = 10000))

## F3 IRT model####
F3_fit <- mirt(F3, model = 1, itemtype = 'graded', SE = TRUE,
               Se.type = 'MHRM',technical = list(NCYCLES = 10000))

## F4 IRT model####
F4_fit <- mirt(F4, model = 1, itemtype = 'graded', SE = TRUE,
               Se.type = 'MHRM',technical = list(NCYCLES = 10000))

## F5 IRT model####
F5_fit <- mirt(F5, model = 1, itemtype = 'graded', SE = TRUE,
               Se.type = 'MHRM',technical = list(NCYCLES = 10000))

# Model Parameters ####
## F1 Model Parameters####
F1_params <- coef(F1_fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F1_items <- data.frame(F1_params$items)
F1l_se <- coef(F1_fit, printSE = TRUE)

## F2 Model Parameters####
F2_params <- coef(F2_fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F2_items <- data.frame(F2_params$items)
F2_se <- coef(F2_fit, printSE = TRUE)

## F3 Model Parameters####
F3_params <- coef(F3_fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F3_items <- data.frame(F3_params$items)
F3_se <- coef(F3_fit, printSE = TRUE)

## F4 Model Parameters####
F4_params <- coef(F4_fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F4_items <- data.frame(F4_params$items)
F4_se <- coef(F4_fit, printSE = TRUE)

## F5 Model Parameters####
F5_params <- coef(F5_fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F5_items <- data.frame(F5_params$items)
F5_se <- coef(F5_fit, printSE = TRUE)

itemparameters <- rbind(F1_items,F2_items, F3_items, F4_items,F5_items)

# Model Fit (degrees of freedom too low to check model fit)
#F1.model <- M2(F1_fit)
#F2.model <- M2(F2_fit)
#F3.model <- M2(F3_fit)
#F4.model <- M2(F4_fit)
#F5.model <- M2(F5_fit)

#Item fit
F1.item.fit<- itemfit(F1_fit, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)

F2.item.fit<- itemfit(F2_fit, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)
F3.item.fit<- itemfit(F3_fit, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)
F4.item.fit<- itemfit(F4_fit, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)

F5.item.fit<- itemfit(F5_fit, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)

#Identifying Missfit items based on RMSEA Value
F1.item_misfits <- subset(F1.item.fit, RMSEA.S_X2 >= .06)
F2.item_misfits <- subset(F2.item.fit, RMSEA.S_X2 >= .06)
F3.item_misfits <- subset(F3.item.fit, RMSEA.S_X2 >= .06)
F4.item_misfits <- subset(F4.item.fit, RMSEA.S_X2 >= .06)
F5.item_misfits <- subset(F5.item.fit, RMSEA.S_X2 >= .06)

# Person Fit ####
## F1 Person fit ####
F1.personfit <- personfit(F1_fit) 
F1.personfit_model_misfits <- subset(F1.personfit, Zh < -2)
rownames(F1.personfit_model_misfits)
nrow(F1.personfit_model_misfits)
hist(F1.personfit$Zh, xlab = "Zh Statistics",main = "F1 Person Fit")
abline(v = -2, lwd = 2, lty = 2)

## F2 Person fit ####
F2.personfit <- personfit(F2_fit) 
F2.personfit_model_misfits <- subset(F2.personfit, Zh < -2)
rownames(F2.personfit_model_misfits)
nrow(F2.personfit_model_misfits)
hist(F2.personfit$Zh, xlab = "Zh Statistics",main = "F1 Person Fit")
abline(v = -2, lwd = 2, lty = 2)

## F3 Person fit ####
F3.personfit <- personfit(F3_fit) 
F3.personfit_model_misfits <- subset(F3.personfit, Zh < -2)
rownames(F3.personfit_model_misfits)
nrow(F3.personfit_model_misfits)
hist(F3.personfit$Zh, xlab = "Zh Statistics",main = "F1 Person Fit")
abline(v = -2, lwd = 2, lty = 2)

## F4 Person fit ####
F4.personfit <- personfit(F4_fit) 
F4.personfit_model_misfits <- subset(F4.personfit, Zh < -2)
rownames(F4.personfit_model_misfits)
nrow(F4.personfit_model_misfits)
hist(F4.personfit$Zh, xlab = "Zh Statistics",main = "F1 Person Fit")
abline(v = -2, lwd = 2, lty = 2)

## F5 Person fit ####
F5.personfit <- personfit(F5_fit) 
F5.personfit_model_misfits <- subset(F5.personfit, Zh < -2)
rownames(F3.personfit_model_misfits)
nrow(F5.personfit_model_misfits)
hist(F5.personfit$Zh, xlab = "Zh Statistics",main = "F1 Person Fit")
abline(v = -2, lwd = 2, lty = 2)

# Plots ####
## F1 Plots ####
F1_OCC <- plot(F1_fit, type = "trace", facet =T, main = "F1: Wearing blue light filters OCC")
F1_info <- plot(F1_fit, type = "infoSE", main = "F1: Wearing blue light filters")
F1_iteminfo <- plot(F1_fit, type = "infotrace", 
                    facet_items = T, main = "F1: Wearing blue light filters")

## F2 Plots ####
F2_OCC <- plot(F2_fit, type = "trace", facet =T, main = "F2: Spending time outdoors OCC")
F2_info <- plot(F2_fit, type = "infoSE", main = "F2: Spending time outdoors")
F2_iteminfo <- plot(F2_fit, type = "infotrace", facet_items = T, main = "F2: Spending time outdoors")

## F3 Plots ####
F3_OCC <-plot(F3_fit, type = "trace", facet =T, main = "F3: Using phone and smart-watch in bed OCC")
F3_info <- plot(F3_fit, type = "infoSE", main = "F3: Using phone and smart-watch in bed")
F3_iteminfo <- plot(F3_fit, type = "infotrace", 
                    facet_items = T, main = "F3: Using phone and smart-watch in bed")

## F4 Plots ####
F4_OCC <-plot(F4_fit, type = "trace", facet =T, main = "F4: Using light before bedtime OCC")
F4_info <- plot(F4_fit, type = "infoSE", main = "F4: Using light before bedtime")
F4_iteminfo <- plot(F4_fit, type = "infotrace", facet_items = T,main = "F4: Using light before bedtime",cex = .8)

## F5 Plots ####
F5_OCC <-plot(F5_fit, type = "trace", facet =T, main = "F5: Using light in the morning and during daytime OCC")
F5_info <- plot(F5_fit, type = "infoSE", main = "F5: Using light in the morning and during daytime")
F5_iteminfo <- plot(F5_fit, type = "infotrace", facet_items = T,main = "F5: Using light in the morning and during daytime")

```

```{r IRTreliability, eval=FALSE, include=FALSE}
#This chunk holds codes for IRT model's conditional  and merginal reliability

#conditional reliability
F1.reliability <- plot(F1_fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="" )
F2.reliability <- plot(F2_fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="" )
F3.reliability <- plot(F3_fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="" )
F4.reliability <- plot(F4_fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="" )
F5.reliability <- plot(F5_fit, type = 'rxx', theta_lim = c(-6, 6), 
      main="" )

##IRT reliability (marginal reliability of scales)
F1.marginal.rel <- marginal_rxx(F1_fit)
F2.marginal.rel <-marginal_rxx(F2_fit)
F3.marginal.rel <-marginal_rxx(F3_fit)
F4.marginal.rel <-marginal_rxx(F4_fit)
F5.marginal.rel <-marginal_rxx(F5_fit)

##Scale characteristic curve
F1.scale <- plot(F1_fit, type = 'score', theta_lim = c(-6, 6), main = "")
F2.scale <- plot(F1_fit, type = 'score', theta_lim = c(-6, 6), main = "")
F3.scale <- plot(F1_fit, type = 'score', theta_lim = c(-6, 6), main = "")
F4.scale <- plot(F1_fit, type = 'score', theta_lim = c(-6, 6), main = "")
F5.scale <- plot(F1_fit, type = 'score', theta_lim = c(-6, 6), main = "")
```

```{r iteminfoggplot3, include=FALSE}
#This chunk holds the code for compiled item info curve for the 1st IRT model

Theta <- matrix(seq(-6,6, by = .1))

##Item Extraction
F1.item1 <- extract.item(F1_fit, 1)#item number
F1.item2 <- extract.item(F1_fit, 2)
F1.item3 <- extract.item(F1_fit, 3)
F2.item1 <- extract.item(F2_fit, 1)
F2.item2 <- extract.item(F2_fit, 2)
F2.item3 <- extract.item(F2_fit, 3)
F2.item4 <- extract.item(F2_fit, 4)
F2.item5 <- extract.item(F2_fit, 5)
F2.item6 <- extract.item(F2_fit, 6)
F3.item1 <- extract.item(F3_fit, 1)
F3.item2 <- extract.item(F3_fit, 2)
F3.item3 <- extract.item(F3_fit, 3)
F3.item4 <- extract.item(F3_fit, 4)
F3.item5 <- extract.item(F3_fit, 5)
F4.item1 <- extract.item(F4_fit, 1)
F4.item2 <- extract.item(F4_fit, 2)
F4.item3 <- extract.item(F4_fit, 3)
F4.item4 <- extract.item(F4_fit, 4)
F5.item1 <- extract.item(F5_fit, 1)
F5.item2 <- extract.item(F5_fit, 2)
F5.item3 <- extract.item(F5_fit, 3)
F5.item4 <- extract.item(F5_fit, 4)
F5.item5 <- extract.item(F5_fit, 5)

##Item Info
F1.item.info.1 <- iteminfo(F1.item1, Theta)
F1.item.info.2 <- iteminfo(F1.item2, Theta)
F1.item.info.3 <- iteminfo(F1.item3, Theta)
F2.item.info.1 <- iteminfo(F2.item1, Theta)
F2.item.info.2 <- iteminfo(F2.item2, Theta)
F2.item.info.3 <- iteminfo(F2.item3, Theta)
F2.item.info.4 <- iteminfo(F2.item4, Theta)
F2.item.info.5 <- iteminfo(F2.item5, Theta)
F2.item.info.6 <- iteminfo(F2.item6, Theta)
F3.item.info.1 <- iteminfo(F3.item1, Theta)
F3.item.info.2 <- iteminfo(F3.item2, Theta)
F3.item.info.3 <- iteminfo(F3.item3, Theta)
F3.item.info.4 <- iteminfo(F3.item4, Theta)
F3.item.info.5 <- iteminfo(F3.item5, Theta)
F4.item.info.1 <- iteminfo(F4.item1, Theta)
F4.item.info.2 <- iteminfo(F4.item2, Theta)
F4.item.info.3 <- iteminfo(F4.item3, Theta)
F4.item.info.4 <- iteminfo(F4.item4, Theta)
F5.item.info.1 <- iteminfo(F5.item1, Theta)
F5.item.info.2 <- iteminfo(F5.item2, Theta)
F5.item.info.3 <- iteminfo(F5.item3, Theta)
F5.item.info.4 <- iteminfo(F5.item4, Theta)
F5.item.info.5 <- iteminfo(F5.item5, Theta)

#Data frame

F1.item.info.data <- as.data.frame(cbind(Theta, F1.item.info.1, 
                                              F1.item.info.2,
                                              F1.item.info.3))

F2.item.info.data <- as.data.frame(cbind(Theta, F2.item.info.1, 
                                              F2.item.info.2,
                                              F2.item.info.3,
                                         F2.item.info.4,
                                         F2.item.info.5,
                                         F2.item.info.6))

F3.item.info.data <- as.data.frame(cbind(Theta, F3.item.info.1, 
                                              F3.item.info.2,
                                              F3.item.info.3,
                                         F3.item.info.4,
                                         F3.item.info.5))

F4.item.info.data <- as.data.frame(cbind(Theta, F4.item.info.1, 
                                              F4.item.info.2,
                                              F4.item.info.3,
                                         F4.item.info.4))
F5.item.info.data <- as.data.frame(cbind(Theta, F5.item.info.1, 
                                              F5.item.info.2,
                                              F5.item.info.3,
                                         F5.item.info.4,
                                         F5.item.info.5))

## GGplot2 item-info
###F1
F1.iteminfo.1 <- ggplot(F1.item.info.data , aes(x=Theta, y=F1.item.info.1)) +
  geom_line()  +apatheme +labs(title = "F1 item16") + geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#E64B3599", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 100, 200))

F1.iteminfo.2 <- ggplot(F1.item.info.data , aes(x=Theta, y=F1.item.info.2)) +
  geom_line()  +apatheme +labs( title = "F1 item36")+geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#E64B3599", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 3, 6))

F1.iteminfo.3 <- ggplot(F1.item.info.data , aes(x=Theta, y=F1.item.info.3)) +
  geom_line()  +apatheme +labs(title = "F1 item17")+geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#E64B3599", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 3, 6), limits = c(0,6))

###F2

F2.iteminfo.1 <- ggplot(F2.item.info.data , aes(x=Theta, y=F2.item.info.1)) +
  geom_line()  +apatheme +labs( title = "F2 item11")+geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#00A08799", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F2.iteminfo.2 <- ggplot(F2.item.info.data , aes(x=Theta, y=F2.item.info.2)) +
  geom_line()  + apatheme +
  labs( title = "F2 item10")+
  geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#00A08799", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F2.iteminfo.3 <- ggplot(F2.item.info.data , aes(x=Theta, y=F2.item.info.3)) +
  geom_line()  +apatheme +
  labs( title = "F2 item12")+
  geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#00A08799", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F2.iteminfo.4 <- ggplot(F2.item.info.data , aes(x=Theta, y=F2.item.info.4)) +
  geom_line()  +apatheme +
  labs( title = "F2 item07")+
  geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#00A08799", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F2.iteminfo.5 <- ggplot(F2.item.info.data , aes(x=Theta, y=F2.item.info.5)) +
  geom_line()  +apatheme +
  labs( title = "F2 item08")+
  geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#00A08799", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F2.iteminfo.6 <- ggplot(F2.item.info.data , aes(x=Theta, y=F2.item.info.6)) +
  geom_line()  +apatheme +
  labs( title = "F2 item09")+
  geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#00A08799", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

###F3
F3.iteminfo.1 <- ggplot(F3.item.info.data , aes(x=Theta, y=F3.item.info.1)) +
  geom_line()  +apatheme +labs( title = "F3 item27")+
 geom_hline(yintercept = .18, colour ="red")+geom_area( fill="3C5488FF", alpha=0.6)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F3.iteminfo.2 <- ggplot(F3.item.info.data , aes(x=Theta, y=F3.item.info.2)) +
  geom_line()  +apatheme +
  labs(title = "F3 item03")+
  geom_hline(yintercept = .18, colour ="red")+geom_area( fill="3C5488FF", alpha=0.6)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F3.iteminfo.3 <- ggplot(F3.item.info.data , aes(x=Theta, y=F3.item.info.3)) +
  geom_line()  +apatheme +
  labs(title = "F3 item40")+
  geom_hline(yintercept = .18, colour ="red")+geom_area( fill="3C5488FF", alpha=0.6)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F3.iteminfo.4 <- ggplot(F3.item.info.data , aes(x=Theta, y=F3.item.info.4)) +
  geom_line()  +apatheme +
  labs(title = "F3 item30")+
  geom_hline(yintercept = .18, colour ="red")+theme(axis.line.x = element_line(color='red'),
        axis.line.y = element_line(color='red'),
        panel.border = element_rect(color = "red",
                                    fill = NA,
                                    size = 1))+geom_area( fill="3C5488FF", alpha=0.6)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F3.iteminfo.5 <- ggplot(F3.item.info.data , aes(x=Theta, y=F3.item.info.5)) +
  geom_line()  +apatheme +
  labs(title = "F3 item41")+
 geom_hline(yintercept = .18, colour ="red")+theme(axis.line.x = element_line(color='red'),
        axis.line.y = element_line(color='red'),
        panel.border = element_rect(color = "red",
                                    fill = NA,
                                    size = 1))+geom_area( fill="3C5488FF", alpha=0.6)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

#F4

F4.iteminfo.1 <- ggplot(F4.item.info.data , aes(x=Theta, y=F4.item.info.1)) +
  geom_line()  +apatheme +labs(title = "F4 item32")+
 geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#4DBBD5FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F4.iteminfo.2 <- ggplot(F4.item.info.data , aes(x=Theta, y=F4.item.info.2)) +
  geom_line()  +apatheme +
  labs(title = "F4 item35")+
  geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#4DBBD5FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F4.iteminfo.3 <- ggplot(F4.item.info.data , aes(x=Theta, y=F4.item.info.3)) +
  geom_line()  +apatheme +
  labs(title = "F4 item38")+
 geom_hline(yintercept = .18, colour ="red")+theme(axis.line.x = element_line(color='red'),
        axis.line.y = element_line(color='red'),
        panel.border = element_rect(color = "red",
                                    fill = NA,
                                    size = 1))+geom_area( fill="#4DBBD5FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1.5, 3), limits = c(0,3))

F4.iteminfo.4 <- ggplot(F4.item.info.data , aes(x=Theta, y=F4.item.info.4)) +
  geom_line()  +apatheme +
  labs(title = "F4 item33")+geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#4DBBD5FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 20, 40))

#F5

F5.iteminfo.1 <- ggplot(F3.item.info.data , aes(x=Theta, y=F5.item.info.1)) +
  geom_line()  +apatheme +labs(title = "F5 item46")+
 geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#7E6148FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1, 2), limits = c(0,2))

F5.iteminfo.2 <- ggplot(F3.item.info.data , aes(x=Theta, y=F5.item.info.2)) +
  geom_line()  +apatheme +
  labs(title = "F5 item45")+
 geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#7E6148FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1, 2), limits = c(0,2))

F5.iteminfo.3 <- ggplot(F3.item.info.data , aes(x=Theta, y=F5.item.info.3)) +
  geom_line()  +apatheme +
  labs(title = "F5 item25")+
  geom_hline(yintercept = .18, colour ="red")+theme(axis.line.x = element_line(color='red'),
        axis.line.y = element_line(color='red'),
        panel.border = element_rect(color = "red",
                                    fill = NA,
                                    size = 1))+geom_area( fill="#7E6148FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1, 2), limits = c(0,2))

F5.iteminfo.4 <- ggplot(F3.item.info.data , aes(x=Theta, y=F5.item.info.4)) +
  geom_line()  +apatheme +
  labs(title = "F5 item04")+
 geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#7E6148FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+scale_y_continuous(breaks=c(0, 1, 2), limits = c(0,2))

F5.iteminfo.5 <- ggplot(F3.item.info.data , aes(x=Theta, y=F5.item.info.5)) +
  geom_line()  +apatheme +
  labs( title = "F5 item01")+
 geom_hline(yintercept = .18, colour ="red")+geom_area( fill="#7E6148FF", alpha=0.4)+xlab(expression(theta)) + 
  ylab(expression(I(theta)))+theme(axis.line.x = element_line(color='red'),
        axis.line.y = element_line(color='red'),
        panel.border = element_rect(color = "red",
                                    fill = NA,
                                    size = 1))+scale_y_continuous(breaks=c(0, 1, 2), limits = c(0,2))

#Panel plots

F1.plots <- cowplot::plot_grid(F1.iteminfo.1,F1.iteminfo.2, F1.iteminfo.3,NULL, NULL, NULL,
                     labels = c("A", "B", "C"),
                     align="h",
                     nrow = 1,
                     label_size = 12,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")

F2.plots <- cowplot::plot_grid(F2.iteminfo.1,F2.iteminfo.2, F2.iteminfo.3,F2.iteminfo.4,F2.iteminfo.5,F2.iteminfo.6,
                     labels = c("D", "E", "F", "G", "H", "I"),
                     align="h",
                     nrow = 1,
                     label_size = 12,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")

F3.plots <- cowplot::plot_grid(F3.iteminfo.1, F3.iteminfo.2, F3.iteminfo.3, F3.iteminfo.4, F3.iteminfo.5,NULL,
                     labels = c("J", "K", "L", "M", "N"),
                     align="h",
                     nrow = 1,
                     label_size = 12,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")

F4.plots <- cowplot::plot_grid(F4.iteminfo.1, F4.iteminfo.2 ,F4.iteminfo.3 ,F4.iteminfo.4,NULL, NULL,
                     labels = c("O", "P", "Q", "R"),
                     align="h",
                     nrow = 1,
                     label_size = 12,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")

F5.plots <- cowplot::plot_grid(F5.iteminfo.1,F5.iteminfo.2,F5.iteminfo.3,F5.iteminfo.4,F5.iteminfo.5,NULL,
                     labels = c("S", "T", "U", "V", "W"),
                     align="h",
                     nrow = 1,
                     label_size = 12,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")

all.factot.itemplots <- cowplot::plot_grid(F1.plots,
                                           F2.plots,
                                           F3.plots,
                                           F4.plots,
                                           F5.plots,

                     align="v",
                     nrow = 5,
                     label_size = 12,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")

ggsave("Figures/S3_Fig.png",all.factot.itemplots, width = 18, height = 12, dpi = 600, bg = "white")

# allitemplots <- cowplot::plot_grid(F1.iteminfo.1,F1.iteminfo.2, F1.iteminfo.3,NULL, NULL,NULL,
#                                    
#                                    F2.iteminfo.1,F2.iteminfo.2, F2.iteminfo.3,F2.iteminfo.4,F2.iteminfo.5,F2.iteminfo.6, 
#                                    
#                                    F3.iteminfo.1, F3.iteminfo.2, F3.iteminfo.3, F3.iteminfo.4, F3.iteminfo.5,   NULL, 
#                                    F4.iteminfo.1, F4.iteminfo.2 ,F4.iteminfo.3 ,F4.iteminfo.4,NULL, NULL,
#                                    F5.iteminfo.1,F5.iteminfo.2,F5.iteminfo.3,F5.iteminfo.4,F5.iteminfo.5,NULL,
#                                    
#                                    
#                      labels = "AUTO",
#                      align="v",
#                      ncol = 6,
#                      label_size = 12,
#                      label_fontfamily = "Arial",
#                      label_fontface = "plain")
# 
# 
# # ggsave("Figures/allitemplots.png",allitemplots, width = 18, height = 12, dpi = 600, bg = "white")
# 
# 
# 
# 
# 
# discarded.items.plot <- cowplot::plot_grid(F3.iteminfo.4,
#                                            F3.iteminfo.5,
#                                            F4.iteminfo.3,
#                                            F5.iteminfo.3,
#                                            F5.iteminfo.5, 
#                                             labels = "AUTO",
#                                    align="v",
#                                    ncol = 5,
#                                    label_size = 12,
#                                    label_fontfamily = "Arial",
#                                    label_fontface = "plain")
#

# ggsave("Figures/discarded.png",discarded.items.plot, width = 18, height = 4, dpi = 600, bg = "white")

```

```{r IRTparameters, eval=FALSE, include=FALSE, results='asis'}
apa_table(itemparameters, caption = "Items discrimination and response category difficulty thresholds of 23 items in  LEBA (n=690)", note = "a = item discrimination parameter; b(1-4) = response category difficulty parameter")
```

```{r shortIRT, include=FALSE}
#This chunk holds codes for our SHORT LEBA IRT with item = 18 (2nd IRT)

#Define model
F1 <- sem.data %>% 
  dplyr::select(item16, item36,  item17)
F2.redu <- sem.data %>% 
  dplyr::select(item11,  item10 , item12, item07, Ritem08, item09)
F3.redu <- sem.data %>% 
 dplyr::select(item27, item03,  item40)

F4.redu <- sem.data %>% 
 dplyr::select(item32, item35,  item33)

F5.redu <- sem.data %>% 
  dplyr::select(item46, item45, item25 )

#fit model
F1_fit <- mirt(F1, model = 1, itemtype = 'graded', 
               SE = TRUE, Se.type = 'MHRM',
               technical = list(NCYCLES = 10000))

F2_fit.redu <- mirt(F2.redu, model = 1, itemtype = 'graded', SE = TRUE,
               Se.type = 'MHRM',technical = list(NCYCLES = 10000))

F3_fit.redu <- mirt(F3.redu, model = 1, itemtype = 'graded', SE = TRUE,
               Se.type = 'MHRM',technical = list(NCYCLES = 10000))

F4_fit.redu <- mirt(F4.redu, model = 1, itemtype = 'graded', SE = TRUE,
               Se.type = 'MHRM',technical = list(NCYCLES = 10000))

F5_fit.redu <- mirt(F5.redu, model = 1, itemtype = 'graded', SE = TRUE,
               Se.type = 'MHRM',technical = list(NCYCLES = 10000))

#Item fit
F1.item.fit<- itemfit(F1_fit, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)

F2.item.fit<- itemfit(F2_fit.redu, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)
F3.item.fit<- itemfit(F3_fit.redu, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)
F4.item.fit<- itemfit(F4_fit.redu, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)

F5.item.fit<- itemfit(F5_fit.redu, fit_stats = c("S_X2", "G2","Zh", "infit"),
                                  impute=10)

#Itemfit short
#Item fit
F1.item.fit<- itemfit(F1_fit, fit_stats = c("S_X2"),
                                  impute=10)

F2.item.fit<- itemfit(F2_fit.redu, fit_stats = c("S_X2"),
                                  impute=10)
F3.item.fit<- itemfit(F3_fit.redu, fit_stats = c("S_X2"),
                                  impute=10)
F4.item.fit<- itemfit(F4_fit.redu, fit_stats = c("S_X2"),
                                  impute=10)

F5.item.fit<- itemfit(F5_fit.redu, fit_stats = c("S_X2"),
                                  impute=10)

itemfit <- rbind(F1.item.fit,F2.item.fit,F3.item.fit,F4.item.fit,F5.item.fit)

colnames(itemfit) <- c("Item","Signed Chi-square", "df", "RMSEA", "p")

#Identifying Missfit items based on RMSEA Value
F1.item_misfits <- subset(F1.item.fit, RMSEA.S_X2 >= .06)
F2.item_misfits <- subset(F2.item.fit, RMSEA.S_X2 >= .06)
F3.item_misfits <- subset(F3.item.fit, RMSEA.S_X2 >= .06)
F4.item_misfits <- subset(F4.item.fit, RMSEA.S_X2 >= .06)
F5.item_misfits <- subset(F5.item.fit, RMSEA.S_X2 >= .06)

# Person Fit ####
## F1 Person fit ####
F1.personfit <- personfit(F1_fit) 
F1.personfit_model_misfits <- subset(F1.personfit, Zh < -2)
rownames(F1.personfit_model_misfits)
nrow(F1.personfit_model_misfits)

F1.person <- ggplot(F1.personfit, aes(x=Zh)) + geom_histogram(binwidth=.5,col=I("black"),fill="#00A08799")+geom_vline(xintercept = -2)+
  labs( y="Number of Participants", x = "Zh Statistics: F1")+apatheme+ ylim(0,600) + xlim(-4,2)

## F2 Person fit ####
F2.personfit <- personfit(F2_fit.redu) 
F2.personfit_model_misfits <- subset(F2.personfit, Zh < -2)
rownames(F2.personfit_model_misfits)
nrow(F2.personfit_model_misfits)

F2.person <- ggplot(F2.personfit, aes(x=Zh)) + geom_histogram(binwidth=.5, col=I("black"),fill="#00A08799")+
  scale_x_continuous(limits = c(-5, 1))+geom_vline(xintercept = -2)+
  labs( y="Number of Participants", x = "Zh Statistics: F2")+apatheme+ylim(0,600) + xlim(-4,2)

## F3 Person fit ####
F3.personfit <- personfit(F3_fit.redu) 
F3.personfit_model_misfits <- subset(F3.personfit, Zh < -2)
rownames(F3.personfit_model_misfits)
nrow(F3.personfit_model_misfits)

F3.person <- ggplot(F3.personfit, aes(x=Zh)) + geom_histogram(binwidth=.5, col=I("black"),fill="#00A08799")+geom_vline(xintercept = -2)+
  labs( y="Number of Participants", x = "Zh Statistics: F3")+apatheme+ylim(0,600) + xlim(-4,2)

## F4 Person fit ####
F4.personfit <- personfit(F4_fit.redu) 
F4.personfit_model_misfits <- subset(F4.personfit, Zh < -2)
rownames(F4.personfit_model_misfits)
nrow(F4.personfit_model_misfits)

F4.person <- ggplot(F4.personfit, aes(x=Zh)) + geom_histogram(binwidth=.5, col=I("black"),fill="#00A08799")+geom_vline(xintercept = -2)+
  labs( y="Number of Participants", x = "Zh Statistics: F4")+apatheme+ylim(0,600) + xlim(-4,2)

## F5 Person fit ####
F5.personfit <- personfit(F5_fit) 
F5.personfit_model_misfits <- subset(F5.personfit, Zh < -2)
rownames(F3.personfit_model_misfits)
nrow(F5.personfit_model_misfits)

F5.person <- ggplot(F5.personfit, aes(x=Zh)) + geom_histogram(binwidth=.5, col=I("black"),fill="#00A08799")+geom_vline(xintercept = -2)+
  labs( y="Number of Participants", x = "Zh Statistics: F5")+apatheme+ylim(0,600) + xlim(-4,2)

# Model Parameters ####
## F1 Model Parameters####
F1_params <- coef(F1_fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F1_items.redu <- data.frame(F1_params$items)
F1l_se <- coef(F1_fit, printSE = TRUE)

## F2 Model Parameters####
F2_params <- coef(F2_fit.redu, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F2_items.redu <- data.frame(F2_params$items)
F2_se <- coef(F2_fit.redu, printSE = TRUE)

## F3 Model Parameters####
F3_params <- coef(F3_fit.redu, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F3_items.redu <- data.frame(F3_params$items)
F3_se <- coef(F3_fit.redu, printSE = TRUE)

## F4 Model Parameters####
F4_params <- coef(F4_fit.redu, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F4_items.redu <- data.frame(F4_params$items)
F4_se <- coef(F4_fit.redu, printSE = TRUE)

## F5 Model Parameters####
F5_params <- coef(F5_fit.redu, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 
F5_items.redu<- data.frame(F5_params$items)
F5_se.redu <- coef(F5_fit.redu, printSE = TRUE)

items.quality.2 <- rbind(F1_items.redu,F2_items.redu,F3_items.redu,F4_items.redu,F5_items.redu )

itemfit2 <- itemfit[,-1]

IRT.details <- as.data.frame(cbind(items.quality.2 ,itemfit2))
IRT.details <- tibble::rownames_to_column(IRT.details, "Items")

#F1.reliability <- plot(F1_fit, type = 'rxx', theta_lim = c(-6, 6),  main="" )
#F2.reliability <- plot(F2_fit.redu, type = 'rxx', theta_lim = c(-6, 6),  main="" )
#F3.reliability <- plot(F3_fit.redu, type = 'rxx', theta_lim = c(-6, 6), main="" )
#F4.reliability <- plot(F4_fit.redu, type = 'rxx', theta_lim = c(-6, 6), main="" )
#F5.reliability <- plot(F5_fit, type = 'rxx', theta_lim = c(-6, 6), main="" )

##IRT reliability (marginal reliability of scales)
#F1.marginal.rel <- marginal_rxx(F1_fit)
#F2.marginal.rel <-marginal_rxx(F2_fit.redu)
#F3.marginal.rel <-marginal_rxx(F3_fit.redu)
#F4.marginal.rel <-marginal_rxx(F4_fit.redu)
#F5.marginal.rel <-marginal_rxx(F5_fit)

# Plots2 ####

#F1_test.info <- plot(F1_fit, type = "infoSE", main = "F1: Wearing blue light filters")

#F2_test.info.redu <- plot(F2_fit.redu, type = "infoSE", main = "F2: Spending time outdoors")

#F3_test.info.redu <- plot(F3_fit.redu, type = "infoSE", main = "F3: Using phone and smart-watch in bed")

#F4_test.info.redu <- plot(F4_fit.redu, type = "infoSE", main = "F4: Using light before bedtime")

#F5_test.info <- plot(F5_fit.redu, type = "infoSE", main = "F5: Using light in the morning and during daytime")

##Scale characteristic curve
#F1.scale <- plot(F1_fit, type = 'score', theta_lim = c(-6, 6), main = "")
#F2.scale <- plot(F2_fit.redu, type = 'score', theta_lim = c(-6, 6), main = "")
#F3.scale <- plot(F3_fit.redu, type = 'score', theta_lim = c(-6, 6), main = "")
#F4.scale <- plot(F4_fit.redu, type = 'score', theta_lim = c(-6, 6), main = "")
#F5.scale <- plot(F5_fit.redu, type = 'score', theta_lim = c(-6, 6), main = "")

```

```{r testinfoggplot2, include=F}

#This chunk holds codes for Test information curve (IRT model 2)

Theta <- matrix(seq(-6,6, by = .1))

F1.T1 <- 0
for(i in 1:ncol(F1)){
  F1.T1 <- F1.T1 + iteminfo(extract.item(F1_fit, i), Theta)
}

F2.T1 <- 0
for(i in 1:ncol(F2.redu)){
  F2.T1 <- F2.T1 + iteminfo(extract.item(F2_fit.redu, i), Theta)
}

F3.T1 <- 0
for(i in 1:ncol(F3.redu)){
  F3.T1 <- F3.T1 + iteminfo(extract.item(F3_fit.redu, i), Theta)
}

F4.T1 <- 0
for(i in 1:ncol(F4.redu)){
  F4.T1 <- F4.T1 + iteminfo(extract.item(F4_fit.redu, i), Theta)
}

F5.T1 <- 0
for(i in 1:ncol(F5.redu)){
  F5.T1 <- F5.T1 + iteminfo(extract.item(F5_fit.redu, i), Theta)
}

F1.test.data <- as.data.frame(cbind(Theta, F1.T1))

F2.test.data <- as.data.frame(cbind(Theta, F2.T1))

F3.test.data <- as.data.frame(cbind(Theta, F3.T1))

F4.test.data <- as.data.frame(cbind(Theta, F4.T1))

F5.test.data <- as.data.frame(cbind(Theta, F5.T1))

Test.data <- as.data.frame(cbind(F1.T1,F2.T1, F3.T1, F4.T1, F5.T1))

Test.data.long <- melt(Test.data)
Test.data.2 <- data.frame(Theta=seq(-6,6, by =.1),Test.data.long)

TIC <- ggplot(Test.data.2,aes(Theta,value))+geom_line(size = 1)+apatheme+
  ggtitle("Test Information Plot")+theme(legend.position = "none")+
  labs(y = "Test Information")+ geom_path()+facet_wrap(~variable,scales = "free") +theme(strip.text.x = element_text(size = 15))+geom_area()+ scale_color_npg()+aes(fill = as.factor(variable))

F1.tic <- ggplot(F1.test.data, aes(x=Theta, y=F1.T1)) +
  geom_line()+apatheme+
  labs(y="Test Information", title = "F1: Wearing blue light filters")+geom_area()+theme(legend.position = "none")+
  aes(fill =  "#E64B3599")+
     scale_fill_identity()+xlab(expression(theta)) + 
  ylab(expression(I(theta)))

F2.tic <- ggplot(F2.test.data, aes(x=Theta, y=F2.T1)) +
  geom_line()+apatheme+
  labs(y="Test Information", title = "F2: Spending time outdoors")+ylim(0,8)+
  geom_area()+aes(fill =  "#4DBBD599")+theme(legend.position = "none")+
     scale_fill_identity()+xlab(expression(theta)) + 
  ylab(expression(I(theta)))

F3.tic <- ggplot(F3.test.data, aes(x=Theta, y=F3.T1)) +
  geom_line()+apatheme+
  labs(y="Test Information", title = "F3: Using phone and smart-watch in bed")+ylim(0,8)+
  geom_area()+aes(fill =  "#00A08799")+theme(legend.position = "none")+
     scale_fill_identity()+xlab(expression(theta)) + 
  ylab(expression(I(theta)))

F4.tic <- ggplot(F4.test.data, aes(x=Theta, y=F4.T1)) +
  geom_line()+apatheme+
  labs(y="Test Information", title = "F4: Using light before bedtime")+
  geom_area()+aes(fill =  "#3C548899")+theme(legend.position = "none")+
     scale_fill_identity()+xlab(expression(theta)) + 
  ylab(expression(I(theta)))

F5.tic <- ggplot(F5.test.data, aes(x=Theta, y=F5.T1)) +
  geom_line()+apatheme+
  labs(y="Test Information", title = "F5:Using light in the morning and during daytime") +ylim(0,8)+geom_area()+aes(fill =  "#B09C8599")+theme(legend.position = "none")+
scale_fill_identity()+xlab(expression(theta)) + 
  ylab(expression(I(theta)))

#panel TIC

TIC.plot <- cowplot::plot_grid(F1.tic,F2.tic, F3.tic,
                                   F4.tic,F5.tic,
                   labels = "AUTO",
                     align="v",
                     ncol = 3,
                     label_size = 15,
                     label_fontfamily = "Arial",
                     label_fontface = "plain")
ggsave("Figures/Figure6.png",TIC.plot, width = 18, height = 12, dpi = 600, bg = "white")

```

```{r Testinfo, fig.env = "sidewaysfigure", fig.cap="Test information curves for the five factors of LEBA: (a) wearing blue light filters (b) spending time outdoors (c) using a phone and smart-watch in bed (d) using light before bedtime (e) using light in the morning and during daytime. Along the x-axis, we plotted the underlying latent trait continuum for each factor. Along the y-axis, we plotted how much information a particular factor is carrying across its latent trait continuum", warning=FALSE, out.width="100%", out.height="120%"}

knitr::include_graphics('Figures/Figure6.png', dpi =600)
```

Subsequently, we treated each factor of the short-LEBA as a unidimensional construct and obtained five test information curves (TICs). As (Figure \@ref(fig:Testinfo)). illustrates, the TICs of the first and fifth factors peaked on the right side of the centre of their latent traits, while the TICs of the other three factors were roughly centred on the respective trait continuum ($\theta$). This points out that the LEBA short-scale estimates the light exposure-related behaviour most precisely near the centre of the trait continuum for the second, third and fourth factors and, in contrast, to the right of the centre for the first and fifth factors [@bakerBasicsItemResponse2017].

Finally, **Supplementary Table 7** summarises the item fit indexes of the LEBA short form. All 18 items yielded RMSEA value $\le$.06, indicating adequate fit to the fitted IRT model. Furthermore, **Supplementary Figure 4** depicts the person fit Zh statistics histogram for the five IRT models. Zh statistics are larger than -2 for most participants, suggesting a good person fit regarding the selected IRT models.

```{r itemfittab, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='asis'}
apa_table(IRT.details , caption ="Item discrimination, response category difficulty thrsholds and fit statistics of the 18 items in short LEBA (n=690)",note = "a = item discrimination parameter; b(1-4) = response category difficulty parameter")
```

```{r perfit, include=F}
#This chunk holds codes for arranging the person fit plot of IRT model 2

personfit <- cowplot::plot_grid(F1.person, F2.person,F3.person,F4.person,F5.person, 
          labels = "AUTO", label_size = 12,align = "v", ncol =2,
                       label_fontfamily = "Arial",
                     label_fontface = "plain")

ggsave("Figures/S4_Fig.png",personfit , width = 10, height = 12, dpi = 600, bg = "white")

```

```{r fscore, include=FALSE}
#This chunk holds codes for score estimation using IRT model 2

LEBA.Total <- sem.data %>% 
  rowwise() %>% 
  mutate(F1.Total = sum(item16, item36,  item17))

LEBA.Total <- LEBA.Total %>% 
  rowwise() %>% 
  mutate(F2.Total = sum(item11,  item10 , item12, item07, Ritem08, item09))

LEBA.Total <- LEBA.Total %>% 
  rowwise() %>% 
  mutate(F3.Total = sum(item27, item03,  item40))

LEBA.Total <- LEBA.Total %>% 
  rowwise() %>% 
  mutate(F4.Total = sum(item32, item35,  item33))

LEBA.Total <- LEBA.Total %>% 
  rowwise() %>% 
  mutate(F5.Total = sum(item46, item45, item25))

F1.estimation <- as.vector(fscores(F1_fit, method = "EAP",
                                           full.scores = TRUE, 
                                           full.scores.SE = F))

F2.estimation <- as.vector(fscores(F2_fit.redu, method = "EAP",
                                           full.scores = TRUE, 
                                           full.scores.SE = F))

F3.estimation <- as.vector(fscores(F3_fit.redu, method = "EAP",
                                           full.scores = TRUE, 
                                           full.scores.SE = F))

F4.estimation <- as.vector(fscores(F4_fit.redu, method = "EAP",
                                           full.scores = TRUE, 
                                           full.scores.SE = F))

F5.estimation <- as.vector(fscores(F5_fit, method = "EAP",
                                           full.scores = TRUE, 
                                           full.scores.SE = F))

Ability.F1 <- data.frame(Rawscore = LEBA.Total$F1.Total,
                      Theta =  F1.estimation)

Ability.F2 <- data.frame(Rawscore = LEBA.Total$F2.Total,
                      Theta =  F2.estimation)

Ability.F3 <- data.frame(Rawscore = LEBA.Total$F3.Total,
                      Theta =  F3.estimation)

Ability.F4 <- data.frame(Rawscore = LEBA.Total$F3.Total,
                      Theta =  F3.estimation)

Ability.F5 <- data.frame(Rawscore = LEBA.Total$F3.Total,
                      Theta =  F3.estimation)

ggplot(Ability.F1, aes(x =Theta, y =Rawscore)) +
  geom_point() +
  geom_smooth(method='lm', se =TRUE, formula= y~x,colour = "red")+
  xlim(-2,2) +
  ylim(0,15)

ggplot(Ability.F2, aes(x =Theta, y =Rawscore)) +
  geom_point() +
  geom_smooth(method='lm', se =TRUE, formula= y~x,colour = "red")+
  xlim(-2,2) +
  ylim(0,15)

ggplot(Ability.F3, aes(x =Theta, y =Rawscore)) +
  geom_point() +
  geom_smooth(method='lm', se =TRUE, formula= y~x,colour = "red")+
  xlim(-2,2) +
  ylim(0,15)

ggplot(Ability.F4, aes(x =Theta, y =Rawscore)) +
  geom_point() +
  geom_smooth(method='lm', se =TRUE, formula= y~x,colour = "red")+
  xlim(-2,2) +
  ylim(0,15)

ggplot(Ability.F5, aes(x =Theta, y =Rawscore)) +
  geom_point() +
  geom_smooth(method='lm', se =TRUE, formula= y~x,colour = "red")+
  xlim(-2,2) +
  ylim(0,15)

Score.correation <- cbind(Ability.F1,Ability.F2, Ability.F3, Ability.F4, Ability.F5)
colnames(Score.correation) <- c("F1.Raw", "F1.Estimated","F2.Raw", "F2.Estimated","F3.Raw", "F3.Estimated","F4.Raw", "F4.Estimated","F5.Raw", "F5.Estimated")

F1.cor <- corr.test(Score.correation$F1.Raw, Score.correation$F1.Estimated)

F2.cor <- corr.test(Score.correation$F2.Raw, Score.correation$F2.Estimated)

F3.cor <- corr.test(Score.correation$F3.Raw, Score.correation$F3.Estimated)

F4.cor <- corr.test(Score.correation$F4.Raw, Score.correation$F4.Estimated)

F5.cor <- corr.test(Score.correation$F5.Raw, Score.correation$F5.Estimated)

#Putting significance star for correlation
F1.p <-F1.cor$p
F1.stars <- ifelse(F1.p < .001, "***"
                   , ifelse(p < .01, "**"
                            , ifelse(p < .05, "*"
                                     , ifelse(p < .10, "+", " "))))

F2.p <-F2.cor$p
F2.stars <- ifelse(F2.p < .001, "**"
                   , ifelse(p < .01, "**"
                            , ifelse(p < .05, "*"
                                     , ifelse(p < .10, "+", " "))))

F3.p <-F3.cor$p
F3.stars <- ifelse(F3.p < .001, "***"
                   , ifelse(p < .01, "**"
                            , ifelse(p < .05, "*"
                                     , ifelse(p < .10, "+", " "))))

F4.p <-F4.cor$p
F4.stars <- ifelse(F4.p < .001, "***"
                   , ifelse(p < .01, "**"
                            , ifelse(p < .05, "*"
                                     , ifelse(p < .10, "+", " "))))

F5.p <-F5.cor$p
F5.stars <- ifelse(F5.p < .001, "***"
                   , ifelse(p < .01, "**"
                            , ifelse(p < .05, "*"
                                     , ifelse(p < .10, "+", " "))))

theta.correlation.matrix <- matrix("NA", nrow =1, ncol = 5)

theta.correlation.matrix[1,1] <- paste(apa(F1.cor$r,2,F), F1.stars,sep="")
theta.correlation.matrix[1,2] <- paste(apa(F2.cor$r,2,F), F2.stars,sep="")
theta.correlation.matrix[1,3] <- paste(apa(F3.cor$r,2,F), F3.stars,sep="")
theta.correlation.matrix[1,4] <- paste(apa(F4.cor$r,2,F), F4.stars,sep="")
theta.correlation.matrix[1,5] <- paste(apa(F5.cor$r,2,F), F5.stars,sep="")

as.data.frame(theta.correlation.matrix)
colnames(theta.correlation.matrix) <- c("F1", "F2", "F3", "F4", "F5")
```

```{r WPdata, warning=F, include=FALSE}
#This chunk holds codes for preparing data for wright-map
library(RColorBrewer)
F1.e <- as.data.frame(F1.estimation)
F2.e <- as.data.frame(F2.estimation)
F3.e <- as.data.frame(F3.estimation)
F4.e <- as.data.frame(F4.estimation)
F5.e <- as.data.frame(F5.estimation)

estimation <- as.data.frame(cbind(F1.e,F2.e,F3.e,F4.e,F5.e))

threshold <- rbind(F1_items.redu[,2:5],
                   F2_items.redu[,2:5],
                   F3_items.redu[,2:5],
                   F4_items.redu[,2:5],
                   F5_items.redu[,2:5])

itemlevelcolors <- matrix(rep(brewer.pal(4, "Set1"), 18), byrow = TRUE, ncol = 4)

```

```{r saveWrightmao, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#This chunk holds codes for saving the wright-map
jpeg("Figures/wm.png", width = 800, height = 750)
WrightMap::wrightMap(estimation, threshold, 
          item.prop = 0.65,

          main.title = "LEBA",
          dim.names = c("F1", "F2", "F3", "F4", "F5"),
          thr.sym.col.fg = itemlevelcolors,
          thr.sym.col.bg = itemlevelcolors,
          vertLines = TRUE,
          show.thr.lab = T,
          dim.color = brewer.pal(5, "Set3"))
dev.off()
```

```{r WP, eval=FALSE, fig.cap="Person Item Map", warning=FALSE, include=FALSE, out.height="120%", out.width="100%"}

knitr::include_graphics('Figures/wm.png')
```

```{r IRTscoreCor, eval=FALSE, include=FALSE, results='asis'}
apa_table (theta.correlation.matrix, caption = "correlation coefficents of obtained scores and estimated latent trait for each factots", note = "* p < 0.05; ** p < 0.01; *** p < 0.001"  )
```

# Discussion

Though there are lots of validated scale to measure light exposure, they don't tell us much about the behavioural aspects pertaining to the light exposure. In that vein we have developed a subjective self-reported scale that can capture light exposure related behaviour in different dimensions.

Authors along with an expert panel generated 48 items and evaluated their quality and relevance and made necessary amendments. A large-scale geographically unconstrained quantitative cross-sectional survey was conducted yielding responses from a large sample (n=`r nrow(data)`). Data collected on the first round was used to explore the latent structure (EFA sample; n=`r nrow(EFA.descriptives)`). Exploratory factor analysis revealed a five factor solution with 25 items. ("Wearing blue light filters", "Spending time outdoors", "Using phone and smart-watch in bed", "Using light before bedtime", and "Using light in the morning and during daytime"). The internal consistency reliability coefficient ordinal alpha ranged between `r apa(min(alpha.tab),2,F)`-`r apa(max(alpha.tab),2,F)`. Except for the fifth factor the internal reliability coefficients were above .70. For the fifth factor it was `r apa((alpha.tab[5,1]),2,F)`. As a rule of thumb reliability coefficient higher than .70 is considered to be satisfactory. However for scales with less than 20 items and at the early developmental stage, a value of .50 is also acceptable [@dall2010developmental; @field2013discovering; @Nunnally1978]. Also all of the factors were highly interpretable in terms of a common theme and contributed essentially towards our aim to capture light exposure related behaviour. This eventually led us to accept the obtained five-factor structure. The total scale exhibited satisfactory internal consistency (McDonald's $\omega_t$=`r apa(omega$omega.tot,2,T)`).

Our CFA analysis (CFA sample; n=`r nrow(CFA.descriptives)`) confirmed the five factor structure we obtained in our EFA thus provided evidence of structural validity.(CFI=`r apa(fit.2[5],2,F)`; TLI=`r apa(fit.2[6],2,F)`; RMSEA=`r apa(fit.2[7],2,F)` [`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]; SRMR=`r apa(fit.2[10],2,F) `). However, in this model, we discarded two items (item 26 $\&$ 37 ) for possible cross-loadings. Also, we accepted one pair of items (item 30 $\&$ 41) to covary their error variance . This item-pair was capturing smart-watch based light exposure behaviour. In retrospect, it seemed plausible that not having  a smart-watch may lead the respondent to answer in a similar fashion for these two items. Thus, we decided to accept the modification to our model. The internal consistency coefficients ordinal alpha for the five factors and the total scale were also satisfactory (Ordinal alpha ranged between `r apa(min(rel.k[2,(1:5)]),2,T)` to `r apa(max(rel.k[2,(1:5)]),2,T)`; McDonald's $\omega_t$=`r apa(rel.k[5,6],2,F)`). On this same data set our measurement invariance analysis indicated the underlying construct of our scale "Light exposure related behaviour" was equivalent across native and non-native English speakers. This indicated the applicability of LEBA on both native and non-native English Speakers. However, this decisions brought up another question to be answered, what was the required grade level to understand the items of our scale? To answer this question we ran secondary analysis where we identified the grade level required using Flesch-Kincaid grade level identification method. Our scale would appear understandable to those who were at least a grade four student and had a age of at least 8.33 years. As a secondary analysis we also assessed the semantic similarity of our scale to the scales recorded in the "Semantic Scale Network" database [@rosenbusch2020semantic]. "LEBA" was  related to  "Sleep Disturbance Scale For Children" (SDSC) [@bruni1996sleep] and "Composite International Diagnostic Interview (CIDI): Insomnia"[@robins1988composite]. Upon inspecting the item contents we found items under "Using phone and smart-watch in bed" and "Using light before bedtime" have semantic overlap with the items of SDSC and CIDI. However, the aim of those scale and ours were different. Items in those two scales were looking into behaviours related to sleep whereas our aim was to capture light exposure related behaviour.

lastly, we developed a short-LEBA (18 items) using IRT analysis. We fitted a graded response model to the combined EFA and CFA sample (n=`r nrow(data)`). We discarded five items (1, 25, 30, 38, & 41) with relatively flat item information curve [I($\theta$) <.20]. Test information curves indicated short form of LEBA is a psychometrically sound measure with adequate coverage of underlying traits to capture different extents of light exposure related behaviours with precision. Item fit index and person fit index for all five fitted model were acceptable providing evidence of validity of our models. Items had diverse item discrimination parameters indicating a good range of discrimination- the ability to differentiate respondents with different levels of the light exposure related behaviour.

Based on all the gathered evidence, we can say LEBA can be used to profile individuals based on their light exposure related behaviours, which can facilitate the development process of individual interventions to promote health. All the five factors of LEBA may identify 'problematic' behaviours that are opposed to good light hygiene.

## Conclusion

We developed a novel self-reported subjective scale-"Light exposure behaviour assessment"(LEBA) to capture light exposure related behaviour. We developed 48 items, judged the relevance and content of the items and conducted a large-scale geographically unrestricted cross-sectional survey. Our EFA gave a five solution with 25 items. A CFA with this 25-item scale again offered a five-factor solution, but this time two more item was discarded. The 23-item "LEBA" was found reliable (internal consistency) and valid (structural validity). A short-form of LEBA was developed using IRT analysis. IRT analysis gave a 18-item scale with a good coverage across the underlying trait continuum. Hence, we could recommend that LEBA can be used to profile individuals according to their light exposure related behaviours.

## Future Direction

Since, LEBA is the first of its kind, estimating convergent validity with other subjective scale was not possible. One way to establish the convergent validity of LEBA is to administer this subjective scale along which some objective measurement scales (e.g. personalised light dosimeter). Though such objective scales do not directly capture light exposure related behaviour, potential insight can be drawn by understanding the behaviour pattern and light exposure. Also, light exposure related behaviours can be dependent upon the socio-economic status as behaviours can be modulated by available scales individual have on their disposal. Our analysis did not consider socio-economic status, as we didn't measure it. Investigating the properties of LEBA while considering different socio-economic status would be a valuable addition.

\newpage

# References

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\endgroup
```